---
title: "DESeq Analysis"
author: "Quynh Tran"
date: "August 2, 2015"
output:
  html_document:
    keep_md: true
---


```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.path='figures/',
               echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(digits=3, scipen=8)
```

This script performs the DESeq analyses on the Juvenile HFD studies.  This script was most recently run on `r date()`.  This uses the input data from the kallisto runs.

```{r data-unique-kallisto, echo=FALSE}
wat_file <- 'Unique_Estimated_Counts_WAT.csv'
quad_file <- 'Unique_Estimated_Counts_Quad.csv'
sample.table_file <- '../sample-mapping.csv'
sample.table <- read.csv(sample.table_file)
sample.table$Barcode <- paste(sample.table$Barcode, "counts.txt", sep="-")
rownames(sample.table) <- as.character(lapply(strsplit(sample.table$Barcode, "-"), "[[",1))

wat_outliers <- c("IonXpress_RNA_040")
poss_outliers <- c("IonXpress_RNA_026", "IonXpress_RNA_027", "IonXpress_RNA_042", "IonXpress_RNA_030", "IonXpress_RNA_041", "IonXpress_RNA_040")

annotation_file <- 'ENSEMBL Annotation - Transcripts.csv'
annotation <- read.csv(annotation_file)

wat.data <- read.csv(wat_file)
quad.data <- read.csv(quad_file)

rownames(wat.data) <- wat.data$Row.names
wat.data <- wat.data[,-1]
wat.counts <- wat.data[,grep("IonXpress",colnames(wat.data))]

rownames(quad.data) <- quad.data$Row.names
quad.data <- quad.data[,-1]
quad.counts <- quad.data[,grep("IonXpress",colnames(quad.data))]

#make them into integers
for (col in colnames(wat.counts)) {
  wat.counts[,col] <- as.integer(wat.counts[,col])
}

for (col in colnames(quad.counts)) {
  quad.counts[,col] <- as.integer(quad.counts[,col])
}
```

The counts data, derived from kallisto and get_highest_kallisto_count_transcript.Rmd are located in `r wat_file` and `r quad_file`.  The sample mapping file is located in `r sample.table_file`.  The annotations are from the file `r annotation_file`.

```{r columns_counts, echo=FALSE}
library(reshape2)
colMax <- function(data) sapply(data, max, na.rm = TRUE)
col_counts_wat <- colSums(wat.counts)
col_max_wat <- colMax(wat.counts)
col_mean_wat <- colMeans(wat.counts)
l <- melt(col_counts_wat, value.name="Total_counts", variable.name="Samples")
m <- melt(col_max_wat, value.name="Max_counts", variable.name="Samples")
mean <- melt(col_mean_wat, value.name="Mean_counts", variable.name="Samples")
c <- cbind(mean,m,l)
c$Mark <- ifelse(rownames(c)%in%poss_outliers, "*", "")
kable(c, caption="Total counts and max count for each eWAT samples")

```
 
  Those with asterisks are those 6 samples that are different from the rest according to PCA.

```{r wat_counts_stats, echo=FALSE}
library(stargazer)
stargazer(wat.counts, type="text", title="WAT Descriptive statistics", digits=1,  out="results/WAT_Descriptive_stats.txt")
stargazer(quad.counts, type="text", title="QUAD Descriptive statistics", digits=1, out="results/QUAD_Descriptive_Stats.txt")

```
##Adipose Tissue

The adipose tissue samples are all eWAT samples

## Adipose Overall Model

This model used only the adipose tissue data, and tested for changes for the interaction between feeding state and diet, then for the main effects of each.

```{r wat-overall-kallisto, echo=FALSE}
wat.counts.noOutliers <- wat.counts[, !colnames(wat.counts)%in%wat_outliers]
library(DESeq2)
#subsettted the sample tables and counts tables
sample.table.wat <- subset(sample.table, Tissue=="eWAT")
sample.table.wat <- sample.table.wat[!rownames(sample.table.wat)%in%wat_outliers,]
sample.table.quad <- subset(sample.table, Tissue=="Quadriceps")

dds.wat.feeding <- DESeqDataSetFromMatrix(colData = sample.table.wat,
                                       countData = wat.counts.noOutliers,  
                                       design= ~ Feeding+Diet+Feeding:Diet)

dds.wat.feeding <- DESeq(dds.wat.feeding, parallel=T)
res.wat.feeding <- results(dds.wat.feeding,
                           parallel=T,
                           name='FeedingFed.DietHigh.Fat.Diet')

res.wat.feeding.main.diet <- results(dds.wat.feeding,
                           parallel=T,
                           name='Diet_High.Fat.Diet_vs_Control.Diet')
res.wat.feeding$log2FoldChange_Diet <- res.wat.feeding.main.diet$log2FoldChange
res.wat.feeding$lfcSE_Diet <- res.wat.feeding.main.diet$lfcSE
res.wat.feeding$pvalue_Diet <- res.wat.feeding.main.diet$pvalue
res.wat.feeding$padj_Diet <- res.wat.feeding.main.diet$padj

res.wat.feeding.main.feeding <- results(dds.wat.feeding,
                           parallel=T,
                           name='Feeding_Fed_vs_Fasted')
res.wat.feeding$log2FoldChange_Feeding <- res.wat.feeding.main.feeding$log2FoldChange
res.wat.feeding$lfcSE_Feeding <- res.wat.feeding.main.feeding$lfcSE
res.wat.feeding$pvalue_Feeding <- res.wat.feeding.main.feeding$pvalue
res.wat.feeding$padj_Feeding <- res.wat.feeding.main.feeding$padj

res.wat.feeding$ensembl_transcript_id <- as.character(lapply(strsplit(rownames(res.wat.feeding), "\\."), "[[", 1))

annotated.res.wat.feeding <- merge(as.data.frame(res.wat.feeding), 
                          annotation, 
                          by='ensembl_transcript_id',
                          all.x=T)
```

Using the combined model, we identified `r table(annotated.res.wat.feeding$padj<0.05)['TRUE']` transcripts with a significant interaction between feeding and diet in WAT. Of which, `r table(annotated.res.wat.feeding$padj<0.05, annotated.res.wat.feeding$log2FoldChange>0)['TRUE', 'TRUE']` genes are up-regulated while `r table(annotated.res.wat.feeding$padj<0.05, annotated.res.wat.feeding$log2FoldChange>0)['TRUE', 'FALSE']` genes are downregulated . 

There are `r table(annotated.res.wat.feeding$padj<0.05, annotated.res.wat.feeding$padj_Diet<0.05)['FALSE','TRUE']` genes had a main effect with respect to Diet and `r table(annotated.res.wat.feeding$padj<0.05, annotated.res.wat.feeding$padj_Feeding<0.05)['FALSE','TRUE']` genes had a main effect with respect to Feeding State.

```{r wat-overall-model-summary, echo=FALSE}
kable(head( annotated.res.wat.feeding[order(annotated.res.wat.feeding$padj_Diet),],5), caption="Top Hits for Effects of Diet")
kable(head( annotated.res.wat.feeding[order(annotated.res.wat.feeding$padj_Feeding),],5), caption="Top Hits for Effects of Feeding")
top20 <- head( annotated.res.wat.feeding[order(annotated.res.wat.feeding$padj),],20)

kable(top20, caption="Top Hits for Interaction between Diet and Feeding")
wat_feeding_results_file <- 'results/DESeq_Combined_Model_WAT_Unique_Kalllisto.csv'
write.csv(annotated.res.wat.feeding, wat_feeding_results_file)
```

```{r detect_outliers, echo=FALSE}
top50 <- head(annotated.res.wat.feeding[order(annotated.res.wat.feeding$padj),],100)
top50_counts <- wat.counts[rownames(wat.counts)%in%top50$ensembl_transcript_id,]
top50_counts$Max <- apply(top50_counts,1,max)
top50_counts$ID <- rownames(top50_counts)
t<-melt(top50_counts, value.name="Counts", variable.name="Samples", id.vars=c("ID","Max"))
t_merge <- merge(t, sample.table.wat, by.x="Samples", by.y="row.names")
t_merge[,c("Samples","ID","SampleID","Feeding", "Diet", "Tissue")] <- lapply(t_merge[,c("Samples","ID","SampleID","Feeding", "Diet", "Tissue")], as.factor)
t2 <- subset(t_merge, Max<400) 
t3 <- subset(t_merge, Max>400 & Max<1000)
t4 <- subset(t_merge, Max>1000 & Max<4000)
t5 <- subset(t_merge, Max>4000 & Max<10000)
t6 <- subset(t_merge, Max>10000)
library(ggplot2)
ggplot(t2, aes(ID, Counts), fill=Diet) +
  geom_boxplot(aes(fill=Diet)) + facet_grid(.~Feeding, scale="free", space="free")+
  geom_text(aes(label=substr(Samples,15,17)),size=3, position = position_jitter(w = 0.4))+ ggtitle("Less than 400 counts") +
  coord_flip()
                  
ggplot(t3, aes(ID, Counts), fill=Diet) +
  geom_boxplot(aes(fill=Diet)) + facet_grid(~Feeding, scale="free", space="free")+ 
  geom_text(aes(label=substr(Samples,15,17)),size=3, angle=-90, position = position_jitter(w = 0.4))+ ggtitle("400 < counts <1000") +
  coord_flip()

ggplot(t4, aes(ID, Counts), fill=Diet) +
  geom_boxplot(aes(fill=Diet)) + facet_grid(~Feeding, scale="free", space="free")+ 
  geom_text(aes(label=substr(Samples,15,17)),size=3, angle=-90, position = position_jitter(w = 0.4))+ ggtitle("1000 < counts <4000") +
  coord_flip()

ggplot(t5, aes(ID, Counts), fill=Diet) +
  geom_boxplot(aes(fill=Diet)) + facet_grid(~Feeding, scale="free", space="free")+ 
  geom_text(aes(label=substr(Samples,15,17)),size=3, angle=-90, position = position_jitter(w = 0.4))+ ggtitle("4000 < counts <10000") +
  coord_flip()

ggplot(t6, aes(ID, Counts), fill=Diet) +
  geom_boxplot(aes(fill=Diet)) + facet_grid(~Feeding, scale="free", space="free")+ 
  geom_text(aes(label=substr(Samples,15,17)),size=3, angle=-90, position = position_jitter(w = 0.4))+ ggtitle("counts > 10,000") +
  coord_flip()
```

```{r deseq-MA-plot-interaction-WAT, fig.cap="MA Plot for Interaction results in WAT"}
plotMA(res.wat.feeding, main="DESeq2-WAT", ylim=c(-4,4))
```

  The 6 samples that look like outliers in the PCA: 026, 027, 030, 040, 041, and 042
```{r deseq-PCA-plot-WAT, echo=FALSE}
library(ggplot2)
dds <- estimateSizeFactors(dds.wat.feeding)
dds <- estimateDispersions(dds)
vsd <- varianceStabilizingTransformation(dds)
names = colData(vsd)$SampleID
col_data = paste(colData(vsd)$Feeding, colData(vsd)$Diet)
p<-plotPCA(vsd, intgroup=c("Feeding","Diet"))
p <- p + geom_text(aes_string(x = "PC1", y = "PC2", label = names), col="black", hjust=-.1, vjust=-.5)
print(p)

par(mfrow=c(1,2))
plot(rank(rowMeans(counts(dds))),
genefilter::rowVars(log2(counts(dds)+1)), main="log2(x+1) transform")
plot(rank(rowMeans(assay(vsd))), genefilter::rowVars(assay(vsd)),
main="VST")

library("pheatmap")
#select <- order(rowMeans(counts(dds,normalized=TRUE)),decreasing=TRUE)[1:20]
selected_rows <- which(res.wat.feeding$padj<0.05)
df <- as.data.frame(colData(dds)[,c("Feeding","Diet")])
#heatmap of significant interaction genes
pheatmap(assay(vsd)[selected_rows,], cluster_rows=TRUE, show_rownames=FALSE, cluster_cols=TRUE, annotation_cof=df)

library(ComplexHeatmap)
library(circlize)
#library(RColorBrewer)
sample.wat.reorder <- sample.table.wat[order(sample.table.wat$Feeding),]
vsd_data <- assay(vsd)
ha_column = HeatmapAnnotation(df = data.frame(Feeding=c(rep("Fasted", 11), rep("Fed", 8)), Diet=c(rep("Control", 6), rep("HFD",5), rep("Control", 4), rep("HFD", 4))), 
    col = list(Diet = c("Control" =  "black", "HFD" = "red"), Feeding=c("Fasted"="Green", "Fed"="Blue")))

Heatmap(vsd_data[selected_rows, rownames(sample.wat.reorder)], column_title="Significant Genes WAT", top_annotation = ha_column, show_row_names=FALSE, cluster_columns=FALSE)

top_genes1 <- head(wat.counts[order(wat.counts$IonXpress_RNA_026, decreasing=TRUE),])
top_genes2 <- head(wat.counts[order(wat.counts$IonXpress_RNA_027, decreasing=TRUE),])
top_genes3 <- head(wat.counts[order(wat.counts$IonXpress_RNA_028, decreasing=TRUE),])
top_genes4 <- head(wat.counts[order(wat.counts$IonXpress_RNA_029, decreasing=TRUE),])
top_genes5 <- head(wat.counts[order(wat.counts$IonXpress_RNA_030, decreasing=TRUE),])
top_genes6 <- head(wat.counts[order(wat.counts$IonXpress_RNA_031, decreasing=TRUE),])
top_genes7 <- head(wat.counts[order(wat.counts$IonXpress_RNA_032, decreasing=TRUE),])
top_genes8 <- head(wat.counts[order(wat.counts$IonXpress_RNA_033, decreasing=TRUE),])
top_genes9 <- head(wat.counts[order(wat.counts$IonXpress_RNA_034, decreasing=TRUE),])
top_genes10 <- head(wat.counts[order(wat.counts$IonXpress_RNA_035, decreasing=TRUE),])
top_genes11 <- head(wat.counts[order(wat.counts$IonXpress_RNA_036, decreasing=TRUE),])
top_genes12 <- head(wat.counts[order(wat.counts$IonXpress_RNA_037, decreasing=TRUE),])
top_genes13 <- head(wat.counts[order(wat.counts$IonXpress_RNA_038, decreasing=TRUE),])
top_genes14 <- head(wat.counts[order(wat.counts$IonXpress_RNA_039, decreasing=TRUE),])
top_genes15 <- head(wat.counts[order(wat.counts$IonXpress_RNA_040, decreasing=TRUE),])
top_genes16 <- head(wat.counts[order(wat.counts$IonXpress_RNA_041, decreasing=TRUE),])
top_genes17 <- head(wat.counts[order(wat.counts$IonXpress_RNA_042, decreasing=TRUE),])
top_genes18 <- head(wat.counts[order(wat.counts$IonXpress_RNA_043, decreasing=TRUE),])
top_genes19 <- head(wat.counts[order(wat.counts$IonXpress_RNA_044, decreasing=TRUE),])

top_exp <- top_genes1
rownames(top_exp) <- c(1,2,3,4,5,6)
top_exp$IonXpress_RNA_026 <- sort(rownames(top_genes1))
top_exp$IonXpress_RNA_027 <- sort(rownames(top_genes2))
top_exp$IonXpress_RNA_028 <- sort(rownames(top_genes3))
top_exp$IonXpress_RNA_029 <- sort(rownames(top_genes4))
top_exp$IonXpress_RNA_030 <- sort(rownames(top_genes5))
top_exp$IonXpress_RNA_031 <- sort(rownames(top_genes6))
top_exp$IonXpress_RNA_032 <- sort(rownames(top_genes7))
top_exp$IonXpress_RNA_033 <- sort(rownames(top_genes8))
top_exp$IonXpress_RNA_034 <- sort(rownames(top_genes9))
top_exp$IonXpress_RNA_035 <- rownames(top_genes10)
top_exp$IonXpress_RNA_036 <- rownames(top_genes11)
top_exp$IonXpress_RNA_037 <- rownames(top_genes12)
top_exp$IonXpress_RNA_038 <- rownames(top_genes13)
top_exp$IonXpress_RNA_039 <- rownames(top_genes14)
top_exp$IonXpress_RNA_040 <- rownames(top_genes15)
top_exp$IonXpress_RNA_041 <- rownames(top_genes16)
top_exp$IonXpress_RNA_042 <- rownames(top_genes17)
top_exp$IonXpress_RNA_043 <- rownames(top_genes18)
top_exp$IonXpress_RNA_044 <- rownames(top_genes19)
top_exp_long <- reshape(top_exp, varying = colnames(top_exp), v.names = "Genes", timevar = "Sample", 
             times = colnames(top_exp), new.row.names = 1:114, direction = "long")

colClasses = c("character", "character")
col.names = c("Gene", "Samples")
table_g <- read.table(text = "",
                 colClasses = colClasses,
                 col.names = col.names)
j=0
for (g in unique(top_exp_long$Genes)){
  sample=""
  j = j+1
  for (i in 1:nrow(top_exp_long)){
    if (g == top_exp_long$Genes[i]){
      sample <- paste(top_exp_long$Sample[i], sample, sep=", ")
    }
    if (i == nrow(top_exp_long)){
        newrow <- data.frame(Gene=g, Samples=sample)
        table_g <- rbind(table_g, newrow)
      }
  }
}
kable(table_g, caption="Top_expressed_genes")
write.csv(table_g, file="top_expressed_genes.csv")
```


# Muscle Tissue

The muscle tissue samples are all quadriceps samples

## Muscle Overall Model

This model used only the quadricep tissue data, and tested for changes for the interaction between feeding state and diet, then for the main effects of each.

```{r quad-overall-kallisto, echo=FALSE}
sample.quad.reorder <- sample.table.quad[order(sample.table.quad$Feeding),]
quad.counts <- quad.counts[,rownames(sample.quad.reorder)]
dds.quad.feeding <- DESeqDataSetFromMatrix(colData = sample.quad.reorder,
                                       countData = quad.counts,  
                                       design= ~ Feeding + Diet + Diet:Feeding)

dds.quad.feeding <- DESeq(dds.quad.feeding, parallel=T)
res.quad.feeding <- results(dds.quad.feeding,
                           parallel=T,
                           name='FeedingFed.DietHigh.Fat.Diet')

res.quad.feeding.main.diet <- results(dds.quad.feeding,
                           parallel=T,
                           name='Diet_High.Fat.Diet_vs_Control.Diet')
res.quad.feeding$log2FoldChange_Diet <- res.quad.feeding.main.diet$log2FoldChange
res.quad.feeding$lfcSE_Diet <- res.quad.feeding.main.diet$lfcSE
res.quad.feeding$pvalue_Diet <- res.quad.feeding.main.diet$pvalue
res.quad.feeding$padj_Diet <- res.quad.feeding.main.diet$padj

res.quad.feeding.main.feeding <- results(dds.quad.feeding,
                           parallel=T,
                           name='Feeding_Fed_vs_Fasted')
res.quad.feeding$log2FoldChange_Feeding <- res.quad.feeding.main.feeding$log2FoldChange
res.quad.feeding$lfcSE_Feeding <- res.quad.feeding.main.feeding$lfcSE
res.quad.feeding$pvalue_Feeding <- res.quad.feeding.main.feeding$pvalue
res.quad.feeding$padj_Feeding <- res.quad.feeding.main.feeding$padj

res.quad.feeding$ensembl_transcript_id <- as.character(lapply(strsplit(rownames(res.quad.feeding), "\\."), "[[", 1))

annotated.res.quad.feeding <- merge(as.data.frame(res.quad.feeding), 
                          annotation, 
                          by='ensembl_transcript_id',
                          all.x=T)
```
  Using the combined model, we identified `r table(annotated.res.quad.feeding$padj<0.05)['TRUE']` transcripts with a significant interaction between feeding and diet in quadriceps.  There are `r table(annotated.res.quad.feeding$padj<0.05, annotated.res.quad.feeding$padj_Feeding<0.05)['FALSE','TRUE']` genes that had a main effect with respect to Feeding State.

```{r quad-overall-model-summary}
kable(head( annotated.res.quad.feeding[order(annotated.res.quad.feeding$padj_Diet),],5), caption="Top Hits for Effects of Diet")
kable(head( annotated.res.quad.feeding[order(annotated.res.quad.feeding$padj_Feeding),],20), caption="Top Hits for Effects of Feeding")
quad_feeding_results_file <- 'results/DESeq_Combined_Model_Quadricep_Unique_Kalllisto.csv'
write.csv(annotated.res.quad.feeding, quad_feeding_results_file)
```


```{r deseq-PCA-heatmap-QUAD, echo=FALSE}
dds_quad <- estimateSizeFactors(dds.quad.feeding)
dds_quad <- estimateDispersions(dds_quad)
vsd_quad <- varianceStabilizingTransformation(dds_quad)
names = colData(vsd_quad)$SampleID
col_data = paste(colData(vsd_quad)$Feeding, colData(vsd_quad)$Diet)
q<-plotPCA(vsd_quad, intgroup=c("Feeding","Diet"))
q <- q + geom_text(aes_string(x = "PC1", y = "PC2", label = names), col="black", hjust=-.1, vjust=-.5)
print(q)
par(mfrow=c(1,2))
plot(rank(rowMeans(counts(dds_quad))),
genefilter::rowVars(log2(counts(dds_quad)+1)), main="QUAD log2(x+1) transform")
plot(rank(rowMeans(assay(vsd_quad))), genefilter::rowVars(assay(vsd_quad)), main="QUAD VST")

#select <- order(rowMeans(counts(dds,normalized=TRUE)),decreasing=TRUE)[1:20]
selected_rows <- which(res.quad.feeding$padj_Feeding<0.05)
df <- as.data.frame(colData(dds_quad)[,c("Feeding","Diet")])
#heatmap of significant interaction genes
pheatmap(assay(vsd_quad)[selected_rows,], cluster_rows=TRUE, show_rownames=FALSE, cluster_cols=TRUE, annotation_cof=df)

ha_col_quad = HeatmapAnnotation(df = data.frame(Feeding=c(rep("Fasted", 12), rep("Fed", 8)), Diet=c(rep("Control", 6), rep("HFD",6), rep("Control", 4), rep("HFD", 4))), 
    col = list(Feeding=c("Fasted"="Green", "Fed"="blue"), Diet = c("Control" =  "black", "HFD" = "red")))

Heatmap(assay(vsd_quad)[selected_rows, rownames(sample.quad.reorder)], column_title="Significant Genes Quadriceps", top_annotation = ha_col_quad, show_row_names=FALSE, cluster_columns=FALSE)

```
# Session Information

```{r session-info}
sessionInfo()
```
