---
title: "HFD_data_pulling"
author: "Quynh Tran"
date: "February 3, 2015"
output: html_document
---

This script is used to pull data from the http://bridgeslab.uthsc.edu/mousedb/ and make a .csv data set for analysis.


```{r pulling data from the database, echo=FALSE}
url.cohort8.group1 <- 'http://bridgeslab.uthsc.edu/mousedb/api/v1/data/?format=json&username=davebridges&api_key=fce3fde2a9a2e5dc9e04c20aad90120a621c50b3&limit=1000&offset=0&animal__cohort=8'

library(rjson)
library(RCurl)
json.cohort8.group1 <- fromJSON(getURL(url.cohort8.group1))

library(plyr)
cohort8.data <- rbind(ldply(lapply(json.cohort8.group1[[2]],function(x) do.call(c,unlist(x,r=F))),as.data.frame))

date_after_fasting <- "2014-10-21"
date_before_fasting <- "2014-10-20"
date_12weeks_MRI <- "2015-01-16"

date_of_interest <- c("2014-10-21", "2014-10-20", "2015-01-16")

sub_cohort8.data <- cohort8.data[cohort8.data$experiment.date %in% date_of_interest,]
#Rename the fasting stage based on the date
sub_cohort8.data$Fasting_stat <- ifelse(sub_cohort8.data$experiment.date=="2014-10-20", "Pre_fasting",ifelse(sub_cohort8.data$experiment.date=="2014-10-21","After_fasting","After_HFD"))

#import body composition data
library(reshape2)
cohort8_weight <-dcast(sub_cohort8.data, animal.MouseID+Fasting_stat~assay.assay, value.var='values')

cohort8_weight$Lean <- as.numeric(cohort8_weight$'Lean Mass')/1000
cohort8_weight$Fat <- as.numeric(cohort8_weight$'Total Fat Mass')/1000
cohort8_weight$Weight <- as.numeric(cohort8_weight$'Body Weight')/1000
#calculate percent body fat
cohort8_weight$Percent.Fat <- cohort8_weight$Fat/cohort8_weight$Weight*100
#make animal id a factor
cohort8_weight$Subject <- as.factor(cohort8_weight$animal.MouseID)
cohort8_weight$Fasting_stat <- as.factor(cohort8_weight$Fasting_stat)
#set Pre_fasting as the reference level
cohort8_weight$Fasting_stat <- relevel(as.factor(cohort8_weight$Fasting_stat), ref='Pre_fasting')

#import CLAMS data
filename <- '../data/raw/Oxymax export of 2014-10-17 10 week C57BL6J.csv'
clams_data <- read.csv(filename)
#remove commented data
clams_data <- clams_data[clams_data$Event.Log=="",]
#data$Subject <- as.factor(data$Subject)
#mark which ones are actually in this experiment
start_fasting_interval <- 128
stop_fasting_interval <- 158
interval_time <- 34
#calculated elapsed time in hours
clams_data$Intervals.Elapsed <- clams_data$Interval - min(clams_data$Interval)
clams_data$Time <- clams_data$Intervals.Elapsed * interval_time/60
subjects <- c(seq(1745,1763,1),seq(1765,1769,1))
cohort8_clams <- clams_data[clams_data$Subject %in% subjects,]
cohort8_clams$Subject <- as.factor(cohort8_clams$Subject)
#determine dark/light cycles
#calculate when the first dark cycle starts
first.dark.cycle <- min(clams_data[clams_data$Light.Dark=='Dark',]$Time)
dark.cycles <- data.frame(
  Day1 = c(Start=first.dark.cycle,End=first.dark.cycle+12),
  Day2 = c(Start=first.dark.cycle+24,End=first.dark.cycle+36),
  Day3 = c(Start=first.dark.cycle+48,End=first.dark.cycle+60),
  Day4 = c(Start=first.dark.cycle+72,End=first.dark.cycle+84),
  Day5 = c(Start=first.dark.cycle+96,End=first.dark.cycle+108),
  Day6 = c(Start=first.dark.cycle+120,End=first.dark.cycle+132),
  Day7 = c(Start=first.dark.cycle+144,End=first.dark.cycle+156),
  Day8 = c(Start=first.dark.cycle+168,End=first.dark.cycle+180),
  Day9 = c(Start=first.dark.cycle+192,End=first.dark.cycle+204),
  Day10 = c(Start=first.dark.cycle+216,End=first.dark.cycle+228)  
)
write.csv(cohort8_clams, file="../data/processed/Cohort8_Oxymax.csv")

```

The data are saved in `r filename`.  This script was most recently updated on `r date()` and is saved in `r getwd()`

# Body Weight Analysis
```{r body-weight, echo=FALSE}
se <- function(x) sd(x)/sqrt(length(x))
library(ggplot2)
library(dplyr)
body_weight_summary <-
  cohort8_weight %>%
  group_by(Fasting_stat) %>%
  summarise(
    Weight = mean(Weight),
    Fat = mean(Fat),
    Lean = mean(Lean),
    Percent.Fat = mean(Percent.Fat))

body_weight_err <-
  cohort8_weight %>%
  group_by(Fasting_stat) %>%
  summarise(
    Weight = se(Weight),
    Fat = se(Fat),
    Lean = se(Lean),
    Percent.Fat = se(Percent.Fat))

body_weight_shapiro <-
  cohort8_weight %>%
  group_by(Fasting_stat) %>%
  summarise(
    Weight = shapiro.test(Weight)$p.value,
    Fat = shapiro.test(Fat)$p.value,
    Lean = shapiro.test(Lean)$p.value,
    Percent.Fat = shapiro.test(Percent.Fat)$p.value)

body_weight_delta <-
  cohort8_weight %>%
  group_by(animal.MouseID) %>%
  mutate(
    Delta_weight=c(0,diff(Weight)))

delta_weight<-body_weight_delta[body_weight_delta$Fasting_stat=="Pre_fasting",c(1,8,9,11)]

library(car)
body.weight.bartlett <- data.frame()
body.weight.bartlett['Weight','pval'] <- bartlett.test(Weight~Fasting_stat, data=cohort8_weight)$p.value
body.weight.bartlett['Fat','pval'] <- bartlett.test(Fat~Fasting_stat, data=cohort8_weight)$p.value
body.weight.bartlett['Lean','pval'] <- bartlett.test(Lean~Fasting_stat, data=cohort8_weight)$p.value
body.weight.bartlett['Percent.Fat','pval'] <- bartlett.test(Percent.Fat~Fasting_stat, data=cohort8_weight)$p.value

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

par(mfrow=c(2,2))
ymax <- max(body_weight_summary$Weight + body_weight_err$Weight)
plot <- barplot(as.numeric(body_weight_summary$Weight),
                beside=T, las=1, main="Body Weight", ylim=c(0,ymax), col=c('blue','green','red'))
superpose.eb(plot,
             as.numeric(body_weight_summary$Weight),
             as.numeric(body_weight_err$Weight))

ymax <- max(body_weight_summary$Fat + body_weight_err$Fat)
plot <- barplot(as.numeric(body_weight_summary$Fat),
                beside=T, las=1, main="Total Fat Mass", ylim=c(0,ymax), col=c('blue','green','red'))

superpose.eb(plot,
             as.numeric(body_weight_summary$Fat),
             as.numeric(body_weight_err$Fat))

ymax <- max(body_weight_summary$Lean + body_weight_err$Lean)
plot <- barplot(as.numeric(body_weight_summary$Lean),
                beside=T, las=1, main="Lean Mass", ylim=c(0,ymax), col=c('blue','green','red'), names.arg=c("Prefasting","AfterFasting","AfterHFD"))
superpose.eb(plot,
             as.numeric(body_weight_summary$Lean),
             as.numeric(body_weight_err$Lean))

ymax <- max(body_weight_summary$Percent.Fat + body_weight_err$Percent.Fat)
plot <- barplot(as.numeric(body_weight_summary$Percent.Fat),
                beside=T, las=1, main="Percent Fat Mass", ylim=c(0,ymax), col=c('blue','green','red'),names.arg=c("Prefasting","AfterFasting","AfterHFD"))
superpose.eb(plot,
             as.numeric(body_weight_summary$Percent.Fat),
             as.numeric(body_weight_err$Percent.Fat))

ggplot(cohort8_weight, aes(x=Fasting_stat, y=Weight, group=Subject, colour=Subject)) + geom_point()+geom_line()+
  xlab("Fasting status")+
  ylab("Weight (g)")
```
## Body Weight Data Statistics

Body weight, total fat mass, lean mass and percet fat mass were all normally distributed (Shapiro-Wilk test; p >`r min(as.matrix(body_weight_shapiro)[2:5])`).  The assumptions of equal variance were not met for each variable (Levene's test p  >`r  min( body.weight.bartlett$pval)`).  Based on this we performed anova with Welch correction on these measurements, as shown below:

```{r body-weight-statistics, results='asis'}
body.weight.oneway <- data.frame()
body.weight.oneway['Weight','pval'] <- oneway.test(Weight~Fasting_stat, data=cohort8_weight)$p.value
body.weight.oneway['Fat','pval'] <- oneway.test(Fat~Fasting_stat, data=cohort8_weight)$p.value
body.weight.oneway['Lean','pval'] <- oneway.test(Lean~Fasting_stat, data=cohort8_weight)$p.value
body.weight.oneway['Percent.Fat','pval'] <- oneway.test(Percent.Fat~Fasting_stat, data=cohort8_weight)$p.value
body.weight.oneway$padj <- p.adjust(body.weight.oneway$pval, method="BH")
library(xtable)
print(xtable(body.weight.oneway, caption="Oneway ANOVA with Welch correction for effects of Fasting and HFD on Body Composition.  P-values are adjusted by the method of Benjamini and Hochberg", digits=8))
```

```{r body-weight-TUKEYHSD, echo=FALSE}
body.weight.tukey <- data.frame()
body.weight.tukey <- TukeyHSD(aov(Weight~Fasting_stat, data=cohort8_weight))$Fasting_stat
print(xtable(body.weight.tukey, caption="Tukey HSD for pairwise comparisons among Fasting stages on Body Weight", digits=8))
body.fat.tukey <- TukeyHSD(aov(Fat~Fasting_stat, data=cohort8_weight))$Fasting_stat
print(xtable(body.fat.tukey, caption="Tukey HSD for pairwise comparisons among Fasting stages on Body Fat", digits=8))
body.lean.tukey<- TukeyHSD(aov(Lean~Fasting_stat, data=cohort8_weight))$Fasting_stat
print(xtable(body.lean.tukey, caption="Tukey HSD for pairwise comparisons among Fasting stages on Lean Mass", digits=8))
percent.fat.tukey<- TukeyHSD(aov(Percent.Fat~Fasting_stat, data=cohort8_weight))$Fasting_stat
print(xtable(percent.fat.tukey, caption="Tukey HSD for pairwise comparisons among Fasting stages on Percent Fat", digits=8))

```

# VO2 Data

```{r vo2-raw-data, echo=FALSE}
#merge the weight and clasm data
annotated_clams_data <- merge(cohort8_weight, cohort8_clams, by='Subject')
annotated_clams_data$VO2 <- annotated_clams_data$Volume.O2*annotated_clams_data$Weight/1000

write.csv(cohort8_clams, file="../data/processed/clams_training_data.csv")

```


````{r correlation_matrix, echo=FALSE}

library(randomForest)
library(miscTools)
library(ggplot2)
library(reshape2)
library(caret)

sub_cohort8_weight <- cohort8_weight[, c(2, 6:10)]
cohort8_weight_wide <- reshape(sub_cohort8_weight, idvar="Subject", timevar="Fasting_stat", 
                               v.names=c("Lean","Fat","Weight","Percent.Fat"), direction="wide")
#calculate the weight loss percentage
cohort8_weight_wide$Percent.WL <- as.numeric((cohort8_weight_wide$Weight.Pre_fasting - cohort8_weight_wide$Weight.After_fasting)/cohort8_weight_wide$Weight.Pre_fasting*100)
cohort8_weight_wide$Percent.WG <- as.numeric((cohort8_weight_wide$Weight.After_HFD - cohort8_weight_wide$Weight.After_fasting)/cohort8_weight_wide$Weight.After_fasting*100)
cohort8_weight_wide$Percent.LL <- as.numeric((cohort8_weight_wide$Lean.Pre_fasting - cohort8_weight_wide$Lean.After_fasting)/cohort8_weight_wide$Lean.Pre_fasting*100)
cohort8_weight_wide$Percent.LG <- as.numeric((cohort8_weight_wide$Lean.After_HFD - cohort8_weight_wide$Lean.After_fasting)/cohort8_weight_wide$Lean.After_fasting*100)
cohort8_weight_wide$Percent.FL <- as.numeric((cohort8_weight_wide$Percent.Fat.Pre_fasting - cohort8_weight_wide$Percent.Fat.After_fasting))
cohort8_weight_wide$Percent.FG <- as.numeric((cohort8_weight_wide$Percent.Fat.After_HFD - cohort8_weight_wide$Percent.Fat.After_fasting))

#save the body composition data
write.csv(cohort8_weight, file="../data/processed/body_compostition_training_data.csv")


panel.cor <- function(x, y, digits=2, cex.cor)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  test <- cor.test(x,y)
  Signif <- ifelse(round(test$p.value,3)<0.001,"p<0.001",paste("p=",round(test$p.value,3)))  
  text(0.5, 0.25, paste("r=",txt))
  text(.5, .75, Signif)
}

panel.smooth<-function (x, y, col = "blue", bg = NA, pch = 18, 
                        cex = 0.8, col.smooth = "red", span = 2/3, iter = 3, ...) 
{
  points(x, y, pch = pch, col = col, bg = bg, cex = cex)
  ok <- is.finite(x) & is.finite(y)
  if (any(ok)) 
    lines(stats::lowess(x[ok], y[ok], f = span, iter = iter), 
          col = col.smooth, ...)
}

panel.hist <- function(x, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col="cyan", ...)
}

pairs(cohort8_weight_wide[,c(14:19)],
          lower.panel=panel.smooth, upper.panel=panel.cor,diag.panel=panel.hist)

```


````{r Using_random_forest, echo=FALSE}
start_fasting_interval <- 128
stop_fasting_interval <- 158

merge_data <- merge(cohort8_weight_wide, annotated_clams_data[,c(1, 11:39)], by='Subject')
merge_data$SumXYAmb <- merge_data$X.Ambulatory+merge_data$Y.Ambulatory

##Calculate the VO2. Before and during fasting, i.e. interval <158, use the Pre_fasting weight to calculate the VO2. After fasting, VO2 will be calculated using the After_fasting weight##
merge_data$VO2.LBM <- ifelse(merge_data$Interval<stop_fasting_interval,merge_data$VO2/merge_data$Lean.Pre_fasting, merge_data$VO2/merge_data$Lean.After_fasting)

merge_data$Fast <- ifelse(merge_data$Interval<start_fasting_interval,"Bf_fasting",ifelse(merge_data$Interval>=stop_fasting_interval, "After_fasting", "During_fasting"))

variables <- c("Percent.WL", "Percent.WG", "Percent.LL", "Percent.LG", "Percent.FL", 
               "Percent.FG")

pairs(merge_data[merge_data$Light.Dark=="Light", c(15, 33, 49, 50)],
          lower.panel=panel.smooth, upper.panel=panel.cor,diag.panel=panel.hist,
          main="LIGHT")

pairs(merge_data[merge_data$Light.Dark=="Dark",c(15, 33, 49, 50)],
          lower.panel=panel.smooth, upper.panel=panel.cor,diag.panel=panel.hist,
        main="DARK")
######Random Forest#######
cols <- c('Subject','Light.Dark', 'Fast', 'Percent.LL','Percent.LG', 'VO2.LBM', 'RER',  'SumXYAmb', 'Percent.WG', 
          "Percent.FL", "Percent.FG")

merge_data$Fast <- as.factor(merge_data$Fast)
model_dat1<-merge_data[,cols]

#model_data <-as.data.frame()
m_data<-
  model_dat1 %>%
  group_by(Subject, Light.Dark, Fast) %>%
  summarise(
    Avg.Percent.LL = mean(Percent.LL),
    Avg.Percent.LG = mean(Percent.LG),
    Avg.Percent.FL = mean(Percent.FL),
    Avg.Percent.FG = mean(Percent.FG),
    RER = mean(RER),
    SumXYAmb = mean(SumXYAmb),
    VO2.LBM = mean(VO2.LBM),
    Percent.WG=mean(Percent.WG))
m_data<-as.data.frame(m_data)
model_data <- m_data[m_data$Fast %in% c("Bf_fasting", "After_fasting"),]

#model_data_wide <- reshape(model_data, timevar="Fast", idvar="Subject", direction="wide", 
 #                               v.names=c( "RER", "VO2.LBM", "SumXYAmb"))

model_data_dark <- model_data[model_data$Light.Dark=="Dark",]
model_data_light <- model_data[model_data$Light.Dark=="Light",]

#make the data into wide format
model_data_dark_wide <- reshape(model_data_dark, timevar="Fast", idvar="Subject", direction="wide", 
                                v.names=c( "RER", "VO2.LBM", "SumXYAmb"))
#calculate the difference in VO2 and RER bf and after fasting
#model_data_dark_wide$VO2.diff <-model_data_dark_wide$VO2.Bf_fasting - model_data_dark_wide$VO2.After_fasting
#model_data_dark_wide$RER.diff <- model_data_dark_wide$RER.Bf_fasting - model_data_dark_wide$RER.After_fasting
#model_data_dark_wide$Amb.diff <- model_data_dark_wide$SumXYAmb.Bf_fasting - model_data_dark_wide$SumXYAmb.After_fasting  

model_data_light_wide <- reshape(model_data_light, timevar="Fast", idvar="Subject", direction="wide", 
                                v.names=c( "RER", "VO2.LBM", "SumXYAmb"))
merge_LD <- merge(model_data_light_wide, model_data_dark_wide[,c(1,8:13)],by="Subject")
#calculate the difference in VO2 and RER bf and after fasting
#model_data_light_wide$VO2.diff <-model_data_light_wide$VO2.Bf_fasting - model_data_light_wide$VO2.After_fasting
#model_data_light_wide$RER.diff <- model_data_light_wide$RER.Bf_fasting - model_data_light_wide$RER.After_fasting
#model_data_light_wide$Amb.diff <- model_data_light_wide$SumXYAmb.Bf_fasting - model_data_light_wide$SumXYAmb.After_fasting
#list of predictors
#x = Light, y=Dark
predictors <- c('Subject', 'Avg.Percent.LG', 
                'Avg.Percent.FL', 'Avg.Percent.FG',
                 'VO2.LBM.Bf_fasting.x',
                'VO2.LBM.After_fasting.y', 
                'RER.After_fasting.x',  
                 'RER.Bf_fasting.y',
                'SumXYAmb.After_fasting.x', 
                'SumXYAmb.Bf_fasting.y',
                'Percent.WG')
predictors1 <- c('Subject', 'Avg.Percent.LL','Avg.Percent.LG', 
                'Avg.Percent.FL', 'Avg.Percent.FG',
                'VO2.LBM.After_fasting.x', 'VO2.LBM.Bf_fasting.x',
                'VO2.LBM.After_fasting.y', 'VO2.LBM.Bf_fasting.y',
                'RER.After_fasting.x',  'RER.Bf_fasting.x',
                 'RER.Bf_fasting.y','RER.Bf_fasting.y',
                'SumXYAmb.After_fasting.x', 'SumXYAmb.Bf_fasting.x',
                'SumXYAmb.Bf_fasting.y','SumXYAmb.Bf_fasting.y')

predictors_dark <- c('Subject', 'Avg.Percent.LL','Avg.Percent.LG', 
                'VO2.Bf_fasting', 
                'RER.After_fasting',  
                'SumXYAmb.Bf_fasting',
                'Percent.WG')
predictors_light <- c('Subject', 'Avg.Percent.LL','Avg.Percent.LG', 
                'VO2.diff', 
                'RER.diff',  
                'SumXYAmb.Bf_fasting',
                'Percent.WG')
#partition the data
idx <- createDataPartition(y=merge_LD$Percent.WG, p=0.75, list=FALSE )
training <- merge_LD[idx,]
testing <- merge_LD[-idx,]
set.seed(300)
#cross validation

tst <- rfcv(merge_LD[, predictors1], merge_LD$Percent.WG, scale="log", step=0.5, cv.fold=3, recursive=T)
pairs(tst$n.var, tst$error.cv)

#making the tuning grid
grid_rf <- expand.grid(.mtry=c(2,4, 8))

m_rf <- train(Percent.WG~., data=training[,predictors], method="rf", metric="Rsquared", tuneGrid=grid_rf)


rf <- randomForest(Percent.WG~., data=training[,predictors], ntree=500, mtry=8)

#look at the importance of the predictors, higer value means more importance
round(importance(rf),2)
r2 <- rSquared(testing$Percent.WG, testing$Percent.WG - predict(rf, testing[,predictors]))

mse <- mean((testing$Percent.WG - predict(rf, testing[,predictors]))^2)

p <- ggplot(aes(x=actual, y=pred),
  data=data.frame(actual=testing$Percent.WG, pred=predict(rf, testing[,predictors])))
p + geom_point() +
  geom_abline(color="red") +
  ggtitle(paste("RandomForest Regression in R r^2=", r2, sep=""))
```

We used random forest to predict the mice weights after HFD. The predictors are Subject, Prefasting weight, After_fasting weight, VO2, RER, Light.Dark, Sum of X and Y ambulatories, and the Fasting status. 