---
title: "HFD_data_pulling"
author: "Quynh Tran"
date: "February 3, 2015"
output: html_document
---

This script is used to pull data from the http://bridgeslab.uthsc.edu/mousedb/ and make a .csv data set for analysis.


```{r pulling data from the database, echo=FALSE}
url.cohort8.group1 <- 'http://bridgeslab.uthsc.edu/mousedb/api/v1/data/?format=json&username=davebridges&api_key=fce3fde2a9a2e5dc9e04c20aad90120a621c50b3&limit=1000&offset=0&animal__cohort=8'

library(rjson)
library(RCurl)
json.cohort8.group1 <- fromJSON(getURL(url.cohort8.group1))

library(plyr)
cohort8.data <- rbind(ldply(lapply(json.cohort8.group1[[2]],function(x) do.call(c,unlist(x,r=F))),as.data.frame))

date_after_fasting <- "2014-10-21"
date_before_fasting <- "2014-10-20"
date_12weeks_MRI <- "2015-01-16"

date_of_interest <- c("2014-10-21", "2014-10-20", "2015-01-16")

sub_cohort8.data <- cohort8.data[cohort8.data$experiment.date %in% date_of_interest,]
#Rename the fasting stage based on the date
sub_cohort8.data$Fasting_stat <- ifelse(sub_cohort8.data$experiment.date=="2014-10-20", "Pre_fasting",ifelse(sub_cohort8.data$experiment.date=="2014-10-21","After_fasting","After_HFD"))

#import body composition data
library(reshape2)
cohort8_weight <-dcast(sub_cohort8.data, animal.MouseID+Fasting_stat~assay.assay, value.var='values')

cohort8_weight$Lean <- as.numeric(cohort8_weight$'Lean Mass')/1000
cohort8_weight$Fat <- as.numeric(cohort8_weight$'Total Fat Mass')/1000
cohort8_weight$Weight <- as.numeric(cohort8_weight$'Body Weight')/1000
#calculate percent body fat
cohort8_weight$Percent.Fat <- cohort8_weight$Fat/cohort8_weight$Weight*100
#make animal id a factor
cohort8_weight$Subject <- as.factor(cohort8_weight$animal.MouseID)
cohort8_weight$Fasting_stat <- as.factor(cohort8_weight$Fasting_stat)
#set Pre_fasting as the reference level
cohort8_weight$Fasting_stat <- relevel(as.factor(cohort8_weight$Fasting_stat), ref='Pre_fasting')

#import CLAMS data
filename <- '../data/raw/Oxymax export of 2014-10-17 10 week C57BL6J.csv'
clams_data <- read.csv(filename)
#remove commented data
clams_data <- clams_data[clams_data$Event.Log=="",]
#data$Subject <- as.factor(data$Subject)
#mark which ones are actually in this experiment
start_fasting_interval <- 128
stop_fasting_interval <- 158
interval_time <- 34
#calculated elapsed time in hours
clams_data$Intervals.Elapsed <- clams_data$Interval - min(clams_data$Interval)
clams_data$Time <- clams_data$Intervals.Elapsed * interval_time/60
subjects <- c(seq(1745,1763,1),seq(1765,1769,1))
cohort8_clams <- clams_data[clams_data$Subject %in% subjects,]
cohort8_clams$Subject <- as.factor(cohort8_clams$Subject)
#determine dark/light cycles
#calculate when the first dark cycle starts
first.dark.cycle <- min(clams_data[clams_data$Light.Dark=='Dark',]$Time)
dark.cycles <- data.frame(
  Day1 = c(Start=first.dark.cycle,End=first.dark.cycle+12),
  Day2 = c(Start=first.dark.cycle+24,End=first.dark.cycle+36),
  Day3 = c(Start=first.dark.cycle+48,End=first.dark.cycle+60),
  Day4 = c(Start=first.dark.cycle+72,End=first.dark.cycle+84),
  Day5 = c(Start=first.dark.cycle+96,End=first.dark.cycle+108),
  Day6 = c(Start=first.dark.cycle+120,End=first.dark.cycle+132),
  Day7 = c(Start=first.dark.cycle+144,End=first.dark.cycle+156),
  Day8 = c(Start=first.dark.cycle+168,End=first.dark.cycle+180),
  Day9 = c(Start=first.dark.cycle+192,End=first.dark.cycle+204),
  Day10 = c(Start=first.dark.cycle+216,End=first.dark.cycle+228)  
)
write.csv(cohort8_clams, file="../data/processed/Cohort8_Oxymax.csv")

```

The data are saved in `r filename`.  This script was most recently updated on `r date()` and is saved in `r getwd()`

# Body Weight Analysis
```{r body-weight, echo=FALSE}
se <- function(x) sd(x)/sqrt(length(x))
library(dplyr)
body_weight_summary <-
  cohort8_weight %>%
  group_by(Fasting_stat) %>%
  summarise(
    Weight = mean(Weight),
    Fat = mean(Fat),
    Lean = mean(Lean),
    Percent.Fat = mean(Percent.Fat))

body_weight_err <-
  cohort8_weight %>%
  group_by(Fasting_stat) %>%
  summarise(
    Weight = se(Weight),
    Fat = se(Fat),
    Lean = se(Lean),
    Percent.Fat = se(Percent.Fat))

body_weight_shapiro <-
  cohort8_weight %>%
  group_by(Fasting_stat) %>%
  summarise(
    Weight = shapiro.test(Weight)$p.value,
    Fat = shapiro.test(Fat)$p.value,
    Lean = shapiro.test(Lean)$p.value,
    Percent.Fat = shapiro.test(Percent.Fat)$p.value)

body_weight_delta <-
  cohort8_weight %>%
  group_by(animal.MouseID) %>%
  mutate(
    Delta_weight=c(0,diff(Weight)))

delta_weight<-body_weight_delta[body_weight_delta$Fasting_stat=="Pre_fasting",c(1,8,9,11)]

library(car)
body.weight.bartlett <- data.frame()
body.weight.bartlett['Weight','pval'] <- bartlett.test(Weight~Fasting_stat, data=cohort8_weight)$p.value
body.weight.bartlett['Fat','pval'] <- bartlett.test(Fat~Fasting_stat, data=cohort8_weight)$p.value
body.weight.bartlett['Lean','pval'] <- bartlett.test(Lean~Fasting_stat, data=cohort8_weight)$p.value
body.weight.bartlett['Percent.Fat','pval'] <- bartlett.test(Percent.Fat~Fasting_stat, data=cohort8_weight)$p.value

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

par(mfrow=c(2,2))
ymax <- max(body_weight_summary$Weight + body_weight_err$Weight)
plot <- barplot(as.numeric(body_weight_summary$Weight),
                beside=T, las=1, main="Body Weight", ylim=c(0,ymax), col=c('blue','green','red'))
superpose.eb(plot,
             as.numeric(body_weight_summary$Weight),
             as.numeric(body_weight_err$Weight))

ymax <- max(body_weight_summary$Fat + body_weight_err$Fat)
plot <- barplot(as.numeric(body_weight_summary$Fat),
                beside=T, las=1, main="Total Fat Mass", ylim=c(0,ymax), col=c('blue','green','red'))

superpose.eb(plot,
             as.numeric(body_weight_summary$Fat),
             as.numeric(body_weight_err$Fat))

ymax <- max(body_weight_summary$Lean + body_weight_err$Lean)
plot <- barplot(as.numeric(body_weight_summary$Lean),
                beside=T, las=1, main="Lean Mass", ylim=c(0,ymax), col=c('blue','green','red'), names.arg=c("Prefasting","AfterFasting","AfterHFD"))
superpose.eb(plot,
             as.numeric(body_weight_summary$Lean),
             as.numeric(body_weight_err$Lean))

ymax <- max(body_weight_summary$Percent.Fat + body_weight_err$Percent.Fat)
plot <- barplot(as.numeric(body_weight_summary$Percent.Fat),
                beside=T, las=1, main="Percent Fat Mass", ylim=c(0,ymax), col=c('blue','green','red'),names.arg=c("Prefasting","AfterFasting","AfterHFD"))
superpose.eb(plot,
             as.numeric(body_weight_summary$Percent.Fat),
             as.numeric(body_weight_err$Percent.Fat))
```
## Body Weight Data Statistics

Body weight, total fat mass, lean mass and percet fat mass were all normally distributed (Shapiro-Wilk test; p >`r min(as.matrix(body_weight_shapiro)[2:5])`).  The assumptions of equal variance were not met for each variable (Levene's test p  >`r  min( body.weight.bartlett$pval)`).  Based on this we performed anova with Welch correction on these measurements, as shown below:

```{r body-weight-statistics, results='asis'}
body.weight.oneway <- data.frame()
body.weight.oneway['Weight','pval'] <- oneway.test(Weight~Fasting_stat, data=cohort8_weight)$p.value
body.weight.oneway['Fat','pval'] <- oneway.test(Fat~Fasting_stat, data=cohort8_weight)$p.value
body.weight.oneway['Lean','pval'] <- oneway.test(Lean~Fasting_stat, data=cohort8_weight)$p.value
body.weight.oneway['Percent.Fat','pval'] <- oneway.test(Percent.Fat~Fasting_stat, data=cohort8_weight)$p.value
body.weight.oneway$padj <- p.adjust(body.weight.oneway$pval, method="BH")
library(xtable)
print(xtable(body.weight.oneway, caption="Oneway ANOVA with Welch correction for effects of Fasting and HFD on Body Composition.  P-values are adjusted by the method of Benjamini and Hochberg", digits=8))
```

```{r body-weight-TUKEYHSD, echo=FALSE}
body.weight.tukey <- data.frame()
body.weight.tukey <- TukeyHSD(aov(Weight~Fasting_stat, data=cohort8_weight))$Fasting_stat
print(xtable(body.weight.tukey, caption="Tukey HSD for pairwise comparisons among Fasting stages on Body Weight", digits=8))
body.fat.tukey <- TukeyHSD(aov(Fat~Fasting_stat, data=cohort8_weight))$Fasting_stat
print(xtable(body.fat.tukey, caption="Tukey HSD for pairwise comparisons among Fasting stages on Body Fat", digits=8))
body.lean.tukey<- TukeyHSD(aov(Lean~Fasting_stat, data=cohort8_weight))$Fasting_stat
print(xtable(body.lean.tukey, caption="Tukey HSD for pairwise comparisons among Fasting stages on Lean Mass", digits=8))
percent.fat.tukey<- TukeyHSD(aov(Percent.Fat~Fasting_stat, data=cohort8_weight))$Fasting_stat
print(xtable(percent.fat.tukey, caption="Tukey HSD for pairwise comparisons among Fasting stages on Percent Fat", digits=8))

```

# VO2 Data

```{r vo2-raw-data}
#merge the weight and clasm data
annotated_clams_data <- merge(cohort8_weight, cohort8_clams, by='Subject')
annotated_clams_data$VO2 <- annotated_clams_data$Volume.O2*annotated_clams_data$Weight/1000

write.csv(cohort8_clams, file="../data/processed/training_data.csv")

```

````{r Using_random_forest, echo=FALSE}
library(randomForest)
library(miscTools)
library(ggplot2)
library(reshape2)
library(caret)

cohort8_weight_wide <- dcast(cohort8_weight, Subject~Fasting_stat,value.var="Weight")
#calculate the weight loss percentage
cohort8_weight_wide$diff <- as.numeric((cohort8_weight_wide$Pre_fasting - cohort8_weight_wide$After_fasting)/cohort8_weight_wide$Pre_fasting*100)

start_fasting_interval <- 128
stop_fasting_interval <- 158

merge_data <- merge(cohort8_weight_wide, cohort8_clams, by='Subject')
merge_data$SumXYAmb <- merge_data$X.Ambulatory+merge_data$Y.Ambulatory

##Calculate the VO2. Before and during fasting, i.e. interval <158, use the Pre_fasting weight to calculate the VO2. After fasting, VO2 will be calculated using the After_fasting weight##
merge_data$VO2 <- ifelse(merge_data$Interval<158,merge_data$Volume.O2*merge_data$Pre_fasting/1000, merge_data$Volume.O2*merge_data$After_fasting/1000)
merge_data$Fast <- ifelse(merge_data$Interval<128,"Bf_fasting",ifelse(merge_data$Interval>=158, "After_fasting", "During_fasting"))
merge_data$diff <- as.numeric((merge_data$Pre_fasting-merge_data$After_fasting)/merge_data$Pre_fasting*100)


ggplot(cohort8_weight, aes(x=Fasting_stat, y=Weight, colour=Subject)) + geom_point()

pearsonR<-cor(as.numeric(cohort8_weight_wide$After_HFD), as.numeric(cohort8_weight_wide$diff))
pearsonR<- format(pearsonR,digits=3)

ggplot(cohort8_weight_wide, aes(x=After_HFD, y=diff)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth(method=lm, se=FALSE)+  # Add linear regression line
  xlab("Weight after HFD (g)")+
  ylab("Percentage weight loss (%)")+
  geom_text(aes(x = 35, y = 10, label = paste("Pearson R = ", pearsonR), parse = TRUE))

######Random Forest#######
cols <- c('Subject','diff','Pre_fasting','After_fasting', 'VO2', 'RER', 'Light.Dark', 'SumXYAmb', 'After_HFD')

merge_data$Fast <- as.factor(merge_data$Fast)
model_dat1<-merge_data[,cols]

#model_data <-as.data.frame()
m_data<-
  model_dat1 %>%
  group_by(Subject, Light.Dark, Fast) %>%
  summarise(
    Pre_fasting = mean(Pre_fasting),
    After_fasting = mean(After_fasting),
    RER = mean(RER),
    SumXYAmb = mean(SumXYAmb),
    After_HFD=mean(After_HFD), 
    VO2 = mean(VO2))
m_data<-as.data.frame(m_data)
model_data <- m_data[m_data$Fast %in% c("Bf_fasting", "After_fasting"),]

model_data_dark <- model_data[model_data$Light.Dark=="Dark",]
model_data_light <- model_data[model_data$Light.Dark=="Light",]

#make the data into wide format
cohort8_weight_wide <- dcast(cohort8_weight, Subject~Fasting_stat,value.var="Weight")
model_data_dark_wide <- reshape(model_data_dark, timevar="Fast", idvar="Subject", direction="wide", v.names=c("RER", "VO2", "SumXYAmb"))

predictors <- c('Subject','Pre_fasting','After_fasting', 'RER.After_fasting', 'RER.Bf_fasting',  'VO2.After_fasting','VO2.Bf_fasting',
                'SumXYAmb.After_fasting', 'SumXYAmb.Bf_fasting', "After_HFD")

#partition the data
idx <- createDataPartition(y=model_data_dark_wide$After_HFD, p=0.75, list=FALSE )
training <- model_data_dark_wide[idx,]
testing <- model_data_dark_wide[-idx,]
set.seed(300)
#making the tuning grid
grid_rf <- expand.grid(.mtry=c(2,4,8))

m_rf <- train(After_HFD~., data=training[,predictors], method="rf", metric="Rsquared", tuneGrid=grid_rf)

rf <- randomForest(After_HFD~., data=training[,predictors], ntree=500, mtry=4)

r2 <- rSquared(testing$After_HFD, testing$After_HFD - predict(rf, testing[,predictors]))

mse <- mean((testing$After_HFD - predict(rf, testing[,predictors]))^2)

p <- ggplot(aes(x=actual, y=pred),
  data=data.frame(actual=testing$After_HFD, pred=predict(rf, testing[,predictors])))
p + geom_point() +
  geom_abline(color="red") +
  ggtitle(paste("RandomForest Regression in R r^2=", r2, sep=""))
```

We used random forest to predict the mice weights after HFD. The predictors are Subject, Prefasting weight, After_fasting weight, VO2, RER, Light.Dark, Sum of X and Y ambulatories, and the Fasting status. 