Body Weight Weight Analysis of Ob/Ob Mice
===========================================


Data Entry
-----------

This was from combined weights over several measurements of Ob/Ob Mice on a C57BL6/J background.  Some animals may appear multiple times in this analysis.  Data is downloaded in csv format from the mousedb website.  This includes only fed weights.

```{r data-import, echo=FALSE}
#these four lines are just to get the data from the database
#raw_filename <- "../data/raw_data.csv"
#download.file("http://bridgeslab.uthsc.edu/mousedb/experiment/data/all.csv",destfile=raw_filename, method='curl')
#data <- read.csv(raw_filename, header=T)
#data <- subset(data, Strain=="ob/ob"&Feeding=="fed"&Assay=="Body Weight")
#fasting_data <- subset(data, Strain=="ob/ob"&Feeding=="fasted"&Assay=="Body Weight")
export_file = "../data/raw/Ob_Ob Raw Weight Data.csv"
export_file_fasting = "../data/raw/Ob_Ob Raw Weight Data Fasting.csv"
#write.csv(data, file=export_file)
#write.csv(fasting_data, file=export_file_fasting)
#import the data

data <- read.csv(export_file, row.names='X')
#recalculated weight in grams
data$Weight <- data$Value/1000

#set default level for genotype to be wt
data$Genotype <- relevel(data$Genotype, ref="+/+")

#set default level for background to be c57bl/6j
data$Background <- relevel(data$Background, ref="C57BL/6J")
```

This data is taken from the file ```r export_file```.  This script was most recently run on ```r date()```.  This includes ```r dim(data)[1]``` fed body weight measurements from ```r dim(data[unique(data$Animal),])[1]``` different mice.

```{r scatter-plot-genotype, dev=c('pdf','png'), warning=FALSE, echo=FALSE}
library(ggplot2)
scatter <- ggplot(data, aes(x=Age, y=Weight)) 
scatter.plot <- scatter + geom_point(aes(color=Genotype))  + scale_x_continuous(limits=c(0, max(data$Age)), name="Age (days)")  + scale_y_continuous(limits=c(0,max(data$Weight)), name="Body Weight (g)") + facet_grid(Background~Gender) + stat_smooth(aes(color=Genotype))
print(scatter.plot)
```

```{r scatter-plot-background, dev=c('png','pdf'), warning=FALSE, echo=FALSE}
scatter <- ggplot(data, aes(x=Age, y=Weight)) 
scatter.plot <- scatter + geom_point(aes(color=Background))  + scale_x_continuous(limits=c(0, max(data$Age)), name="Age (days)")  + scale_y_continuous(limits=c(0,max(data$Weight)), name="Body Weight (g)") + facet_grid(Genotype~Gender) + stat_smooth(aes(color=Background))
print(scatter.plot)
```

```{r summary, echo=FALSE}
library(plyr)
#adult is defined as >100 days but <400 days
summary.adult <- ddply(subset(data, Age > 100 & Age <400 & Feeding=='fed'), c("Background","Genotype"), summarise, 
  mean = mean(Weight),
  se = sd(Weight)/sqrt(length(Weight)),
  sd = sd(Weight),
  rel.sd = sd(Weight)/mean(Weight)*100,
  number = length(Weight),
  age = mean(Age))
```

Analysis by Age
------------------

We segregated mice into groups of of adult mice (100-400 days) and calculated some basic statistics.

```{r data-tables,echo=FALSE,results='asis',message=FALSE}
require(xtable)
print(xtable(summary.adult, caption = "Summarized Data for Adult Mice", label = "tab:adult-summary"), include.rownames=F, type='html')
```

```{r statistics, echo=FALSE, message=FALSE}
require(lme4)
nlme.fit <- lmer(log(Weight)~ Age + Background + Genotype + Age:Background + Age:Genotype + (1|Animal), data=data)
nlme.fit.nolog <- lmer(Weight ~ Age + Background + Genotype + Age:Background + Age:Genotype + (1|Animal), data=data)

#models with factors removed
nlme.fit.no.age <- lmer(log(Weight) ~ Background + Genotype + (1|Animal), data=data)
nlme.fit.no.geno <- lmer(log(Weight) ~ Age + Background + Age:Background + (1|Animal), data=data)
nlme.fit.no.background <- lmer(log(Weight) ~ Age + Genotype + Age:Genotype + (1|Animal), data=data)

#models with interactions
nlme.fit.background.age <- lmer(log(Weight) ~ Age + Background + Genotype + Age:Genotype + (1|Animal), data=data)
nlme.fit.geno.age <- lmer(log(Weight)~ Age + Background + Genotype + Age:Background + (1|Animal), data=data)
#summary(nlme.fit)

#this is for the post-hoc hypothesis testing
require(multcomp)
ghlt.nlme.fit.genotype <- glht(nlme.fit, linfct=mcp(Genotype="Dunnett"))
#summary(ghlt.nlme.fit.genotype)
ghlt.nlme.fit.background <- glht(nlme.fit, linfct=mcp(Background="Dunnett"))
#summary(ghlt.nlme.fit.background)
```

To summarise these data we generated a mixed linear model comparing the logarithm of the body weight to the genotype, background and age.  The logaritm fit the data much better than a linear effect (p=`r anova(nlme.fit.nolog, nlme.fit)$"Pr(>Chisq)"[2]`).  Each of the other terms were also significant contributors to the model with removal of age (p=`r round(anova(nlme.fit.no.age, nlme.fit)$"Pr(>Chisq)"[2],84)```), genotype (p=`r round(anova(nlme.fit.no.geno, nlme.fit)$"Pr(>Chisq)"[2],30)`) or background strain (p=`r round(anova(nlme.fit.no.background, nlme.fit)$"Pr(>Chisq)"[2],19)`) all having significantly worse models.  There was a significant interaction between genotype and age (p=`r round(anova(nlme.fit.geno.age, nlme.fit)$"Pr(>Chisq)"[2],13)`) and background and age (p=`r round(anova(nlme.fit.background.age, nlme.fit)$"Pr(>Chisq)"[2],4)`), so those terms are included as interactors.  Therefore the final model is:

```{r formula}
formula(nlme.fit)
```

The residuals for this model are:

```{r residuals-plot, echo=FALSE, dev=c('pdf','png')}
par(mfrow=c(1,2))
plot(data$Weight, resid(nlme.fit), col=data$Genotype, 
     pch=19, cex=0.5,
     ylab="Residuals (g)",
     xlab="Actual Weight (g)"
     )
abline(a=0,b=0)
legend("topright", levels(data$Genotype), pch=19, col=palette()[1:length(levels(data$Genotype))], bty="n", cex=0.5)

plot(data$Age, resid(nlme.fit), col=data$Genotype, 
     pch=19, cex=0.5,
     ylab="Residuals (g)",
     xlab="Actual Age (d)"
     )
abline(a=0,b=0)
legend("topright", levels(data$Genotype), pch=19, col=palette()[1:length(levels(data$Genotype))], bty="n", cex=0.5)
```

Variation in Weight Gain
--------------------------

To examine the variation in weight gain between wild-type and knockout mice we generated a density plot and graphed both the Standard Deviation and Relative Standard Deviation.  This is all done for mice >100 days old. 

```{r ob-variation, echo=FALSE}
c57.f.test <- var.test(
  subset(data, Age>110&Age<130&Genotype=='-/-'&Background=='C57BL/6J'&Feeding=='fed')$Weight,
  subset(data, Age>110&Age<130&Genotype=='+/+'&Background=='C57BL/6J'&Feeding=='fed')$Weight)

btbr.f.test <- var.test(
  subset(data, Age>110&Age<130&Genotype=='-/-'&Background=='BTBR'&Feeding=='fed')$Weight,
  subset(data, Age>110&Age<130&Genotype=='+/+'&Background=='BTBR'&Feeding=='fed')$Weight)
```

 To test whether these differences are statistically significant we did a F-test comparing the standard deviations of these groups.  Neither difference was statistically significant (p=`r round(c57.f.test$p.value, 3)` for C57BL/6J and p=`r round(btbr.f.test$p.value, 3)` for BTBR).

```{r ob-density,dev=c('png','pdf'),echo=FALSE}
par(mfrow=c(1,2))
with(subset(data, Age>110&Age<130&Genotype=='-/-'&Background=='C57BL/6J'&Feeding=='fed'), plot(density(Weight), 
                  xlim=c(0, max(Weight)),
                  xlab="Weight (g)",
                  ylim = c(0,0.20),
                  las=1, main="C57BL/6J"))
lines(density(subset(data, Age>110&Age<130&Genotype=='+/+'&Background=='C57BL/6J'&Feeding=='fed')$Weight), lty=2)
legend('topleft', c("Wild-Type","ob/ob"), bty="n", lty=c(2,1))

with(subset(data, Age>110&Age<130&Genotype=='-/-'&Background=='BTBR'&Feeding=='fed'), plot(density(Weight), 
                  xlim=c(0, max(Weight)),
                  xlab="Weight (g)",
                  ylim = c(0,0.20),
                  las=1, main="BTBR"))
lines(density(subset(data, Age>110&Age<130&Genotype=='+/+'&Background=='BTBR'&Feeding=='fed')$Weight), lty=2)
```

```{r ob-density-btbr,dev=c('png','pdf'),echo=FALSE}
with(subset(data, Age>110&Age<130&Genotype=='-/-'&Background=='BTBR'&Feeding=='fed'), plot(density(Weight), 
                  xlim=c(0, max(Weight)),
                  xlab="Weight (g)",
                  ylim = c(0,0.20),
                  las=1, main="BTBR"))
lines(density(subset(data, Age>110&Age<130&Genotype=='+/+'&Background=='BTBR'&Feeding=='fed')$Weight), lty=2)
```

```{r ob-density-c57bl6j,dev=c('png','pdf'),echo=FALSE}
with(subset(data, Age>110&Age<130&Genotype=='-/-'&Background=='C57BL/6J'&Feeding=='fed'), plot(density(Weight), 
                  xlim=c(0, max(Weight)),
                  xlab="Weight (g)",
                  ylim = c(0,0.20),
                  las=1, main="C57BL/6J"))
lines(density(subset(data, Age>110&Age<130&Genotype=='+/+'&Background=='C57BL/6J'&Feeding=='fed')$Weight), lty=2)
legend('topleft', c("Wild-Type","ob/ob"), bty="n", lty=c(2,1))
```

```{r ob-sd,dev=c('png','pdf'),echo=FALSE}
useful.summary.adult <- droplevels(subset(summary.adult, Genotype %in% c('+/+','-/-')))
barplot(useful.summary.adult$sd, las=1, 
        names.arg=useful.summary.adult$Genotype,
        ylab="Standard Deviation (g)",
        col=useful.summary.adult$Background)
legend('topright', levels(useful.summary.adult$Background), fill=palette()[1:2], bty='n')
```

```{r ob-rel-sd,dev=c('png','pdf'),echo=FALSE}
barplot(useful.summary.adult$rel.sd, las=1, 
        names.arg=useful.summary.adult$Genotype,
        ylab="Relative Standard Deviation (%)",
        col=useful.summary.adult$Background)
legend('topright', levels(useful.summary.adult$Background), fill=palette()[1:2], bty='n')
```

Fasting Response in ob/ob Mice
-------------------------------

Fasting data is taken from `r export_file_fasting`.

```{r fasting-data, echo=FALSE}
fasted.data <- read.csv(export_file_fasting, row.names='X')
fasted.data$Weight <- fasted.data$Value/1000
combined.data <- merge(fasted.data, data, by="Animal", all.y=F)
#subsetted so each row is a pair of fed/fasted separated by one day
fasted.combined.data <- combined.data[(combined.data$Age.x - combined.data$Age.y == 1),]
fasted.combined.data$Loss <- fasted.combined.data$Weight.y - fasted.combined.data$Weight.x
fasted.combined.data$Loss.pct <- fasted.combined.data$Loss/fasted.combined.data$Weight.y*100

fasted.combined.data <- droplevels(subset(fasted.combined.data, Genotype.x %in% c('+/+','-/-')))

shapiro.tests.fasting <- data.frame(
  row.names=c("BTBR +/+", "BTBR -/-", "C57BL/6J +/+", "C57BL/6J -/-"),
  Day70 = rep(NA, 4),
  Day120 = rep(NA, 4))

shapiro.tests.fasting['C57BL/6J +/+','Day70'] <- with(subset(fasted.combined.data, Genotype.x=="+/+"&Background.x=="C57BL/6J"&Age.x>68&Age.x<73), shapiro.test(Loss)$p.value)
shapiro.tests.fasting['C57BL/6J -/-','Day70'] <- with(subset(fasted.combined.data, Genotype.x=="-/-"&Background.x=="C57BL/6J"&Age.x>68&Age.x<73), shapiro.test(Loss)$p.value)
shapiro.tests.fasting['C57BL/6J +/+','Day120'] <- with(subset(fasted.combined.data, Genotype.x=="+/+"&Background.x=="C57BL/6J"&Age.x>115&Age.x<125), shapiro.test(Loss)$p.value)
shapiro.tests.fasting['C57BL/6J -/-','Day120'] <- with(subset(fasted.combined.data, Genotype.x=="-/-"&Background.x=="C57BL/6J"&Age.x>115&Age.x<125), shapiro.test(Loss)$p.value)
shapiro.tests.fasting['BTBR +/+','Day120'] <- with(subset(fasted.combined.data, Genotype.x=="+/+"&Background.x=="BTBR"&Age.x>115&Age.x<125), shapiro.test(Loss)$p.value)
shapiro.tests.fasting['BTBR -/-','Day120'] <- with(subset(fasted.combined.data, Genotype.x=="-/-"&Background.x=="BTBR"&Age.x>115&Age.x<125), shapiro.test(Loss)$p.value)
```

To determine whether the fasting data has a normal distribution we did Shapiro-Wilk tests on each of the groups:

```{r shapiro-tests, results='asis', echo=FALSE}
print(xtable(shapiro.tests.fasting), type='html')
```

First we tested whether there was an effect of genotype on the fasting response at approximately 10 weeks of age.  The t-test between wild-type and knockouts having different absolute fasting induced weight loss is `r round(t.test(Loss ~ Genotype.x, data = subset(fasted.combined.data, Age.x>68&Age.x<73&Genotype.x!='N.D.'&Genotype.x!='-/+'))$p.value,4)` and for percent weight loss is `r round(t.test(Loss.pct ~ Genotype.x, data = subset(fasted.combined.data, Age.x>68&Age.x<73&Genotype.x!='N.D.'&Genotype.x!='-/+'))$p.value,4)`.  These data are graphed below

```{r weight-loss,dev=c('png','pdf'),echo=FALSE, message=FALSE}
require(reshape2)
loss.mean <- as.matrix(dcast(subset(fasted.combined.data, Age.x>68&Age.x<73&Genotype.x!='N.D.'), Background.x~Genotype.x, mean, value.var="Loss")[2:3])
se <- function(x) sd(x, na.rm=T)/sqrt(length(x))
loss.se <- as.matrix(dcast(subset(fasted.combined.data, Age.x>68&Age.x<73&Genotype.x!='N.D.'), Background.x~Genotype.x, se, value.var="Loss")[2:3])

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
length = length, ...)

ymax <- max(loss.mean+loss.se)
plot <- barplot(loss.mean, las=1, ylab="Fasting Weight Loss (g)", ylim=c(0,ymax), main="10 Week Old Mice (C57BL/6J Mice Only)")
superpose.eb(plot, loss.mean, loss.se)
```

```{r weight-loss-pct,dev=c('png','pdf'),echo=FALSE}
loss.mean <- as.matrix(dcast(subset(fasted.combined.data, Age.x>68&Age.x<73&Genotype.x!='N.D.'), Background.x~Genotype.x, mean, value.var="Loss.pct")[2:3])
loss.se <- as.matrix(dcast(subset(fasted.combined.data, Age.x>68&Age.x<73&Genotype.x!='N.D.'), Background.x~Genotype.x, se, value.var="Loss.pct")[2:3])

ymax <- max(loss.mean+loss.se)
plot <- barplot(loss.mean, las=1, ylab="Fasting Weight Loss (%)", ylim=c(0,ymax), main="10 Week Old Mice (C57BL/6J Mice Only)")
superpose.eb(plot, loss.mean, loss.se)
```

At approximately 120 days of age for C57BL/6J, the t-test between wild-type and knockouts having different absolute fasting induced weight loss is `r round(t.test(Loss ~ Genotype.x, data = subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'&Genotype.x!='-/+'&Background.x=="C57BL/6J"))$p.value,4)` and for percent weight loss is `r round(t.test(Loss.pct ~ Genotype.x, data = subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'&Genotype.x!='-/+'&Background.x=="C57BL/6J"))$p.value,4)`.  For BTBR, the t-test between wild-type and knockouts having different absolute fasting induced weight loss is `r round(t.test(Loss ~ Genotype.x, data = subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'&Genotype.x!='-/+'&Background.x=="BTBR"))$p.value,4)` and for percent weight loss is `r round(t.test(Loss.pct ~ Genotype.x, data = subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'&Genotype.x!='-/+'&Background.x=="BTBR"))$p.value,4)`.  These data are graphed below.

```{r weight-loss-120d,dev=c('png','pdf'),echo=FALSE}
loss.mean <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, mean, value.var="Loss")[1:2,2:3])
loss.se <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, se, value.var="Loss")[1:2,2:3])

ymax <- max(loss.mean+loss.se, na.rm=T)
plot <- barplot(loss.mean, las=1, ylab="Fasting Weight Loss (g)", 
                beside=2, ylim=c(0,ymax), main="120 Day Old Mice")
superpose.eb(plot, loss.mean, loss.se)
legend("topright",levels(fasted.combined.data$Genotype.x), fill=grey.colors(2), bty="n")
```

```{r weight-loss-120d-c57bl6j,dev=c('png','pdf'),echo=FALSE}
loss.mean <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, mean, value.var="Loss")[1:2,3])
loss.se <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, se, value.var="Loss")[1:2,3])

ymax <- max(loss.mean+loss.se, na.rm=T)
plot <- barplot(as.vector(loss.mean), las=1, ylab="Fasting Weight Loss (g)", beside=T,names.arg=levels(fasted.combined.data$Genotype.x),
                ylim=c(0,ymax), main="120 Day Old Mice (C57BL/6J Only)")
superpose.eb(plot, loss.mean, loss.se)
```

```{r weight-loss-120d-btbr,dev=c('png','pdf'),echo=FALSE}
loss.mean <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, mean, value.var="Loss")[1:2,2])
loss.se <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, se, value.var="Loss")[1:2,2])

ymax <- max(loss.mean+loss.se, na.rm=T)
plot <- barplot(as.vector(loss.mean), las=1, ylab="Fasting Weight Loss (g)", beside=T,names.arg=levels(fasted.combined.data$Genotype.x),
                ylim=c(0,ymax), main="120 Day Old Mice (BTBR Only)")
superpose.eb(plot, loss.mean, loss.se)
```

```{r weight-loss-pct-120d,dev=c('png','pdf'),echo=FALSE}
loss.mean <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, mean, value.var="Loss.pct")[1:2,2:3])
loss.se <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, se, value.var="Loss.pct")[1:2,2:3])

ymax <- max(loss.mean+loss.se)
plot <- barplot(loss.mean, beside=T, las=1, ylab="Fasting Weight Loss (%)", ylim=c(0,ymax), main="120 Day Old Mice")
superpose.eb(plot, loss.mean, loss.se)
legend("topright",levels(fasted.combined.data$Genotype.x), fill=grey.colors(2), bty="n")
```

```{r weight-loss-pct-120d-c57bl6j,dev=c('png','pdf'),echo=FALSE}
loss.mean <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, mean, value.var="Loss.pct")[1:2,2])
loss.se <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, se, value.var="Loss.pct")[1:2,2])

ymax <- max(loss.mean+loss.se)
plot <- barplot(as.vector(loss.mean), beside=T, las=1, 
                names.arg=levels(fasted.combined.data$Genotype.x), ylab="Fasting Weight Loss (%)", ylim=c(0,ymax), main="120 Day Old Mice (C57BL/6J Only")
superpose.eb(plot, loss.mean, loss.se)
```

```{r weight-loss-pct-120d-btbr,dev=c('png','pdf'),echo=FALSE}
loss.mean <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, mean, value.var="Loss.pct")[1:2,2])
loss.se <- as.matrix(dcast(subset(fasted.combined.data, Age.x>115&Age.x<125&Genotype.x!='N.D.'), Genotype.x~Background.x, se, value.var="Loss.pct")[1:2,2])

ymax <- max(loss.mean+loss.se)
plot <- barplot(as.vector(loss.mean), beside=T, las=1, 
                names.arg=levels(fasted.combined.data$Genotype.x), ylab="Fasting Weight Loss (%)", ylim=c(0,ymax), main="120 Day Old Mice (BTBR Only)")
superpose.eb(plot, loss.mean, loss.se)
```

Session Information
---------------------
  
```{r session-info}
sessionInfo()
```

