Analysis of The Predictive Value of Pre-Diet Serum Hormones from High Fat Diet Cohort 6
========================================================================================

Summary
--------

Serum samples were taken from mice which were fasted for 16h.  These mice were 10 weeks old, prior to being placed on diets.  

```{r data-entry, echo=FALSE}
#set palette
library(RColorBrewer)
palette(brewer.pal(3, "Set2")[c(3,2)])
data_file <- "../data/raw/Pre-Diet-Multiplex.csv"
#this uses the median data
data <- read.csv(data_file, skip=46, nrows=(130-46))
useful.data <- data[,4:11] #removed extra rows
rownames(useful.data) <- data$Sample


data.blanked <- sweep(as.matrix(useful.data), 2, as.numeric(useful.data['Blank',]), "-")
data.standards <- data.blanked[grepl("Standard", rownames(data.blanked)),]
data.samples <- data.blanked[!(grepl("Standard", rownames(data.blanked))),]
#remove unknown samples
data.samples <- data.samples[!(grepl("Unknown", rownames(data.samples))),]
#remove blanks
data.samples <- data.samples[!(grepl("Blank", rownames(data.samples))),]
#remove values under 200 which were the samples in cohort 5
data.cohort.5 <- as.data.frame(data.samples[32:73,])
rownames(data.cohort.5) <- as.integer(as.character((rownames(data.cohort.5))))

#merged with mapping file
mapping_file <- "../data/raw/mapping.csv" #includes Treatment and Feeding Status
mapping <- read.csv(mapping_file, row.names="X")
merged_data <- merge(data.cohort.5,mapping, by=0)

#import weight data
#start <- "http://bridgeslab.uthsc.edu/mousedb/experiment/100/csv"
#end <- "http://bridgeslab.uthsc.edu/mousedb/experiment/126/csv"
pre.diet.datafile <- '../data/raw/Cohort_6_PreDiet.csv'
post.diet.datafile <- '../data/raw/Cohort_6_PostDiet.csv'
#download.file(start, pre.diet.datafile)
#download.file(end, post.diet.datafile)

start.data <- read.csv(pre.diet.datafile)
end.data <- read.csv(post.diet.datafile)


combined.data <- merge(start.data, end.data, by="Animal")[,c('Animal','Values.x','Values.y')]
rownames(combined.data) <- as.numeric(lapply(strsplit(as.character(combined.data$Animal),"#"), "[[", 2))
colnames(combined.data) <- c('Animal', 'Before', 'After')

combined.data$Gain <- combined.data$After - combined.data$Before
combined.data$Gain.pct <- combined.data$Gain/combined.data$Before*100
combined.data$Gain <- combined.data$Gain/1000 #converted to grams

merged_weight_data <- merge(combined.data, merged_data, by.x=0,by.y="Row.names")
```


The raw data is in a file named `r data_file` and are mapped against the file in `r mapping_file`.  The mouse level data is shown below.  This analysis looked at `r table(merged_weight_data$Treatment)[1]` Control Diet mice and `r table(merged_weight_data$Treatment)[2]` High Fat Diet mice.  The pre-diet weights are shown in `r pre.diet.datafile` and the post-diet weights are shown in `r post.diet.datafile`.

```{r data-tables,echo=FALSE,results='asis'}
require(xtable)
print(xtable(merged_weight_data[order(merged_weight_data$Treatment, -merged_weight_data$Gain.pct),], caption = "Mouse Level Data", label = "tab:animal-data"), include.rownames=F, type="html")
```

Statistics
-----------

```{r statistics-pct-weight, echo=FALSE, warning=FALSE}
normality.tests <- data.frame(row.names=colnames(merged_weight_data)[3:14],
                              'Control Diet' = rep(NA, 12),
                              'High Fat Diet' = rep(NA, 12))

for (measure in rownames(normality.tests)) {
  normality.tests[measure,] <- tapply(merged_weight_data[,measure], merged_weight_data$Treatment, FUN=function(x) shapiro.test(x)$p.value)
}
normality.tests$Normal <- apply(normality.tests, 1, min) > 0.05

hormones.tested <- colnames(merged_weight_data)[7:14]
merged_weight_data <- subset(merged_weight_data, Row.names!='192')
aov.model.null <- lm(Gain.pct~Treatment, data=merged_weight_data)

summary.lm <- data.frame(row.names = hormones.tested,
    p.value = rep(NA, length(hormones.tested)),
    R.HFD = rep(NA, length(hormones.tested)),
    R.CD= rep(NA, length(hormones.tested)),
    R2.HFD = rep(NA, length(hormones.tested)),
    R2.CD= rep(NA, length(hormones.tested)),
    p.HFD = rep(NA, length(hormones.tested)),
    p.CD= rep(NA, length(hormones.tested)))

summary.lm['Resistin', ] <- c(
  add1(aov.model.null, ~ Treatment * Resistin, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Resistin,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Resistin,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Resistin,Gain.pct, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Resistin,Gain.pct, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Resistin,Gain.pct, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Resistin,Gain.pct, method='spearman'))$p.value)

summary.lm['GIP', ] <- c(
  add1(aov.model.null, ~ Treatment * GIP, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GIP,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GIP,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GIP,Gain.pct, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GIP,Gain.pct, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GIP,Gain.pct, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GIP,Gain.pct, method='spearman'))$p.value)

summary.lm['PAI1', ] <- c(
  add1(aov.model.null, ~ Treatment * PAI1, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(PAI1,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(PAI1,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(PAI1,Gain.pct, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(PAI1,Gain.pct, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(PAI1,Gain.pct, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(PAI1,Gain.pct, method='spearman'))$p.value)

summary.lm['GLP1', ] <- c(
  add1(aov.model.null, ~ Treatment * GLP1, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GLP1,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GLP1,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GLP1,Gain.pct, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GLP1,Gain.pct, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GLP1,Gain.pct, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GLP1,Gain.pct, method='spearman'))$p.value)

summary.lm['Glucagon', ] <- c(
  add1(aov.model.null, ~ Treatment * Glucagon, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Glucagon,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Glucagon,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Glucagon,Gain.pct, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Glucagon,Gain.pct, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Glucagon,Gain.pct, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Glucagon,Gain.pct, method='spearman'))$p.value)

summary.lm['Ghrelin', ] <- c(
  add1(aov.model.null, ~ Treatment * Ghrelin, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Ghrelin,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Ghrelin,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Ghrelin,Gain.pct, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Ghrelin,Gain.pct, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Ghrelin,Gain.pct, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Ghrelin,Gain.pct, method='spearman'))$p.value)

summary.lm['Leptin', ] <- c(
  add1(aov.model.null, ~ Treatment * Leptin, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Leptin,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Leptin,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Leptin,Gain.pct, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Leptin,Gain.pct, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Leptin,Gain.pct, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Leptin,Gain.pct, method='spearman'))$p.value)

summary.lm['Insulin', ] <- c(
  add1(aov.model.null, ~ Treatment * Insulin, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Insulin,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Insulin,Gain.pct, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Insulin,Gain.pct, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Insulin,Gain.pct, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Insulin,Gain.pct, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Insulin,Gain.pct, method='spearman'))$p.value)

summary.lm.abs <- data.frame(row.names = hormones.tested,
    p.value = rep(NA, length(hormones.tested)),
    R.HFD = rep(NA, length(hormones.tested)),
    R.CD= rep(NA, length(hormones.tested)),
    R2.HFD = rep(NA, length(hormones.tested)),
    R2.CD= rep(NA, length(hormones.tested)),
    p.HFD = rep(NA, length(hormones.tested)),
    p.CD= rep(NA, length(hormones.tested)))

aov.model.null.abs <- lm(Gain~Treatment, data=merged_weight_data)

summary.lm.abs['Resistin', ] <- c(
  add1(aov.model.null.abs, ~ Treatment * Resistin, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Resistin,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Resistin,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Resistin,Gain, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Resistin,Gain, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Resistin,Gain, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Resistin,Gain, method='spearman'))$p.value)

summary.lm.abs['GIP', ] <- c(
  add1(aov.model.null.abs, ~ Treatment * GIP, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GIP,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GIP,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GIP,Gain, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GIP,Gain, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GIP,Gain, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GIP,Gain, method='spearman'))$p.value)

summary.lm.abs['PAI1', ] <- c(
  add1(aov.model.null.abs, ~ Treatment * PAI1, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(PAI1,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(PAI1,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(PAI1,Gain, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(PAI1,Gain, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(PAI1,Gain, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(PAI1,Gain, method='spearman'))$p.value)

summary.lm.abs['GLP1', ] <- c(
  add1(aov.model.null.abs, ~ Treatment * GLP1, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GLP1,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GLP1,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GLP1,Gain, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GLP1,Gain, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(GLP1,Gain, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(GLP1,Gain, method='spearman'))$p.value)

summary.lm.abs['Glucagon', ] <- c(
  add1(aov.model.null.abs, ~ Treatment * Glucagon, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Glucagon,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Glucagon,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Glucagon,Gain, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Glucagon,Gain, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Glucagon,Gain, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Glucagon,Gain, method='spearman'))$p.value)

summary.lm.abs['Ghrelin', ] <- c(
  add1(aov.model.null.abs, ~ Treatment * Ghrelin, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Ghrelin,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Ghrelin,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Ghrelin,Gain, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Ghrelin,Gain, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Ghrelin,Gain, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Ghrelin,Gain, method='spearman'))$p.value)

summary.lm.abs['Leptin', ] <- c(
  add1(aov.model.null.abs, ~ Treatment * Leptin, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Leptin,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Leptin,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Leptin,Gain, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Leptin,Gain, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Leptin,Gain, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Leptin,Gain, method='spearman'))$p.value)

summary.lm.abs['Insulin', ] <- c(
  add1(aov.model.null.abs, ~ Treatment * Insulin, test="F")[2,6],
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Insulin,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Insulin,Gain, method='spearman'))$estimate,
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Insulin,Gain, method='spearman'))$estimate^2,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Insulin,Gain, method='spearman'))$estimate^2,  
  with(subset(merged_weight_data, Treatment=="High Fat Diet"), cor.test(Insulin,Gain, method='spearman'))$p.value,
  with(subset(merged_weight_data, Treatment=="Control Diet"), cor.test(Insulin,Gain, method='spearman'))$p.value)

absolute.summary.file <- '../data/processed/Pre-Diet Hormone Correlations - Absolute Weight Gain.csv'
percent.summary.file <- '../data/processed/Pre-Diet Hormone Correlations - Percent Weight Gain.csv'
write.csv(summary.lm.abs, absolute.summary.file)
write.csv(summary.lm, percent.summary.file)
```
We first tested each of the measured factors for whether the values fit a normal distribution.  These data are shown below:

```{r normality-tables,echo=FALSE,results='asis'}
print(xtable(normality.tests, caption = "Shapiro-Wilk tests for each measured variable", label = "tab:normality-tests", digits=5), type='html')
```

### Comparason of factors to percent weight gain

We tested for correlations between each hormone and the percent body weight gained on a high fat diet.  The null model with just the treatment term predicting the weight gain is described below.  This null model had an adjusted R<sup>2</sup> of `r summary(aov.model.null)$adj.r.squared`.

```{r correlation-tables,echo=FALSE,results='asis'}
print(xtable(summary.lm[order(summary.lm$p.value),], caption = "Model and Correlation Summary.  Linear Model Results for Effects Pre-Diet Hormones on Percent Weight Gain (Corrected for Diet) is shown in p.value, correlation coefficients baed on Spearman's Rank Order Test for each treatment is also shown", label = "tab:hormone-summary", digits=3), type='html')
```

We tested whether including the pre-diet hormone levels will improve the model in which the pre-diet hormones are not taken into account.  We did this allowing for an interaction between the diet and the hormone levels.  The p-value is the results of a F-test against whether the model is improved by including that hormone data.  No hormones significantly improved the prediction of this model.  These data are summarized in the file `r percent.summary.file`.

```{r anova-tables,echo=FALSE,results='asis'}
print(xtable(aov.model.null, caption = "ANOVA Determining the Effects of Diet on Eventual Weight Gain", label = "tab:model-null", digits=3), type='html')
```

```{r scatter-plot,dev=c('png','pdf'),echo=FALSE}
par(mfrow=c(2,4))

with(merged_weight_data, plot(Resistin, Gain.pct, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain.pct)), xlim=c(0,max(Resistin)),
         ylab="Percent Weight Gain",
         xlab="Hormone Levels",
         main="Resistin"))
abline(lm(Gain.pct~Resistin, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain.pct~Resistin, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(GIP, Gain.pct, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain.pct)), xlim=c(0,max(GIP)),
         ylab="Percent Weight Gain",
         xlab="Hormone Levels",
         main="GIP"))
abline(lm(Gain.pct~GIP, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain.pct~GIP, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(PAI1, Gain.pct, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain.pct)), xlim=c(0,max(PAI1)),
         ylab="Percent Weight Gain",
         xlab="Hormone Levels",
         main="PAI1"))
abline(lm(Gain.pct~PAI1, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain.pct~PAI1, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(GLP1, Gain.pct, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain.pct)), xlim=c(0,max(GLP1)),
         ylab="Percent Weight Gain",
         xlab="Hormone Levels",
         main="GLP1"))
abline(lm(Gain.pct~GLP1, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain.pct~GLP1, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(Glucagon, Gain.pct, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain.pct)), xlim=c(0,max(Glucagon)),
         ylab="Percent Weight Gain",
         xlab="Hormone Levels",
         main="Glucagon"))
abline(lm(Gain.pct~Glucagon, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain.pct~Glucagon, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(Ghrelin, Gain.pct, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain.pct)), xlim=c(0,max(Ghrelin)),
         ylab="Percent Weight Gain",
         xlab="Hormone Levels",
         main="Ghrelin"))
abline(lm(Gain.pct~Ghrelin, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain.pct~Ghrelin, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(Leptin, Gain.pct, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain.pct)), xlim=c(0,max(Leptin)),
         ylab="Percent Weight Gain",
         xlab="Hormone Levels",
         main="Leptin"))
abline(lm(Gain.pct~Leptin, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain.pct~Leptin, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(Insulin, Gain.pct, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain.pct)), xlim=c(0,max(Insulin)),
         ylab="Percent Weight Gain",
         xlab="Hormone Levels",
         main="Insulin"))
abline(lm(Gain.pct~Insulin, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain.pct~Insulin, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])
```

### Comparason of factors to absolute weight gain

Next, we tested for correlations between each hormone and the absolute body weight gained on a high fat diet.  The null model with just the treatment term predicting the weight gain is described below.  This null model had an adjusted R<sup>2</sup> of `r summary(aov.model.null.abs)$adj.r.squared`.

```{r correlation-tables-abs,echo=FALSE,results='asis'}
print(xtable(summary.lm.abs[order(summary.lm.abs$p.value),], caption = "Model and Correlation Summary.  Linear Model Results for Effects Pre-Diet Hormones on Absolute Weight Gain (Corrected for Diet) is shown in p.value, correlation coefficients baed on Spearman's Rank Order Test for each treatment is also shown", label = "tab:hormone-summary-abs", digits=3), type='html')
```

We tested whether including the pre-diet hormone levels will improve the model in which the pre-diet hormones are not taken into account.  We did this allowing for an interaction between the diet and the hormone levels.  The p-value is the results of a F-test against whether the model is improved by including that hormone data.  No hormones significantly improved the prediction of this model.  These data are summarized in the file `r absolute.summary.file`.

```{r anova-tables-abs,echo=FALSE,results='asis'}
print(xtable(aov.model.null.abs, caption = "ANOVA Determining the Effects of Diet on Eventual Weight Gain", label = "tab:model-null", digits=3), type='html')
```

```{r scatter-plot-abs,dev=c('png','pdf'),echo=FALSE}
par(mfrow=c(2,4))

with(merged_weight_data, plot(Resistin, Gain, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain)), xlim=c(0,max(Resistin)),
         ylab="Weight Gain (g)",
         xlab="Hormone Levels",
         main="Resistin"))
abline(lm(Gain~Resistin, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain~Resistin, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(GIP, Gain, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain)), xlim=c(0,max(GIP)),
         ylab="Weight Gain (g)",
         xlab="Hormone Levels",
         main="GIP"))
abline(lm(Gain~GIP, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain~GIP, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(PAI1, Gain, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain)), xlim=c(0,max(PAI1)),
         ylab="Weight Gain (g)",
         xlab="Hormone Levels",
         main="PAI1"))
abline(lm(Gain~PAI1, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain~PAI1, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(GLP1, Gain, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain)), xlim=c(0,max(GLP1)),
         ylab="Weight Gain (g)",
         xlab="Hormone Levels",
         main="GLP1"))
abline(lm(Gain~GLP1, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain~GLP1, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(Glucagon, Gain, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain)), xlim=c(0,max(Glucagon)),
         ylab="Weight Gain (g)",
         xlab="Hormone Levels",
         main="Glucagon"))
abline(lm(Gain~Glucagon, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain~Glucagon, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(Ghrelin, Gain, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain)), xlim=c(0,max(Ghrelin)),
         ylab="Weight Gain (g)",
         xlab="Hormone Levels",
         main="Ghrelin"))
abline(lm(Gain~Ghrelin, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain~Ghrelin, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(Leptin, Gain, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain)), xlim=c(0,max(Leptin)),
         ylab="Weight Gain (g)",
         xlab="Hormone Levels",
         main="Leptin"))
abline(lm(Gain~Leptin, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain~Leptin, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])

with(merged_weight_data, plot(Insulin, Gain, 
         col=Treatment, pch=19, las=2,frame.plot=F,
         ylim=c(0,max(Gain)), xlim=c(0,max(Insulin)),
         ylab="Weight Gain (g)",
         xlab="Hormone Levels",
         main="Insulin"))
abline(lm(Gain~Insulin, subset(merged_weight_data, Treatment=="High Fat Diet")), col=palette()[2])
abline(lm(Gain~Insulin, subset(merged_weight_data, Treatment=="Control Diet")), col=palette()[1])
```

```{r bibliography, include=F, echo=F}
require(bibtex)
write.bib(c(names(sessionInfo()$otherPkgs),'base','stats','utils'), file='analysis-references.bib')
```

Session Information
--------------------
```{r sessionInfo, echo=F}
sessionInfo()
```