---
title: "Compare_cohorts"
author: "Quynh Tran"
date: "November 4, 2015"
output: pdf_document
---

This script was most recently run on `r date()`. This script is to compare the mice body weight across different cohorts. The script also explores the relationship between initial weight, weight loss and the weights after HFD.

```{r process_cohort5_6_data, message=FALSE, echo=FALSE, warning=FALSE}
previous_cohort_file <- "../data/processed/fasting_weights_file.csv"
prev_cohort_data <- read.csv(previous_cohort_file)

#only get HFD mice from previous cohorts
prev.data <- prev_cohort_data[prev_cohort_data$Treatment=="High Fat Diet",]
prev.data <- prev.data[!is.na(prev.data$X),]
prev.data$Subject <- substring(prev.data$Animal, 18,20)

prev.long <- reshape(prev.data, 
  varying = c("Early.Fed", "Early.Fasted", "Late.Fed", "Late.Fasted"), 
  v.names = "Weight",
  timevar = "Fasting_stat", 
  times = c("Early.Fed", "Early.Fasted", "Late.Fed", "Late.Fasted"), 
  new.row.names = 1:1000,
  direction = "long")
prev.long$Fasting_stat <- ifelse(prev.long$Fasting_stat=="Early.Fed", "Pre_fasted", ifelse(prev.long$Fasting_stat=="Early.Fasted", "Fasted", ifelse(prev.long$Fasting_stat=="Late.Fed", "After_HFD", prev.long$Fasting_stat)))


sub.prev.long <- prev.long[prev.long$Fasting_stat%in% c("Pre_fasted", "Fasted", "After_HFD"),]
sub.prev.long <- sub.prev.long[!is.na(sub.prev.long$X),]

#create a data set for cohort 5 and 6 that has the same structure as cohort 8 and 9
cohort5.6 <- data.frame(animal.MouseID=sub.prev.long$Subject, Fasting_stat=sub.prev.long$Fasting_stat, Lean = numeric(nrow(sub.prev.long)), Fat=numeric(nrow(sub.prev.long)), Weight=sub.prev.long$Weight, Percent.Fat=numeric(nrow(sub.prev.long)), Subject=sub.prev.long$Subject, Cohort=substring(sub.prev.long$Cohort, 8,9))


```
Weight data from cohort 5 and 6 were read in from `r previous_cohort_file`.

```{r process_cohort7, message=FALSE, echo=FALSE, warning=FALSE}
cohort7_file <- "../data/raw/Cohort_7_Raw_Data.csv"
cohort8_file <- "../data/raw/Cohort_8_Raw_Data.csv"

cohort7.data <- read.csv(cohort7_file, row.names="X")
cohort8.data <- read.csv(cohort8_file, row.names="X")

starting_date <- '2014-10-17'
date_after_fasting <- '2014-10-21'
date_before_fasting <- '2014-10-20'
date_12weeks_MRI <- '2015-01-16'


date_of_interest <- c('2014-10-21', '2014-10-20', '2015-01-16')

sub_cohort7.data <- cohort7.data[cohort7.data$experiment.date %in% date_of_interest,]
#Rename the fasting stage based on the date
sub_cohort7.data$Fasting_stat <- ifelse(sub_cohort7.data$experiment.date=="2014-10-20", "Pre_fasted",ifelse(sub_cohort7.data$experiment.date=="2014-10-21","Fasted","After_HFD"))
sub_cohort7.data$animal.MouseID <- as.factor(sub_cohort7.data$animal.MouseID)

#import body composition data
library(reshape2)
library(plyr)
#there should be Lean Mass for this record, but some how it was labeled "Total Fat Mass"
sub_cohort7.data$assay.assay[which(sub_cohort7.data$assay.assay=="Total Fat Mass" & sub_cohort7.data$values=="22200")] <- "Lean Mass"

cohort7_weight <-dcast(sub_cohort7.data, animal.MouseID+Fasting_stat~assay.assay, value.var='values')

cohort7_weight$Lean <- as.numeric(cohort7_weight$'Lean Mass')/1000
cohort7_weight$Fat <- as.numeric(cohort7_weight$'Total Fat Mass')/1000
cohort7_weight$Weight <- as.numeric(cohort7_weight$'Body Weight')/1000
#calculate percent body fat
cohort7_weight$Percent.Fat <- cohort7_weight$Fat/cohort7_weight$Weight*100
#make animal id a factor
cohort7_weight$Subject <- as.factor(cohort7_weight$animal.MouseID)
cohort7_weight$Fasting_stat <- as.factor(cohort7_weight$Fasting_stat)
#set Pre_fasting as the reference level
cohort7_weight$Fasting_stat <- relevel(as.factor(cohort7_weight$Fasting_stat), ref='Pre_fasted')
cohort7_weight$Cohort <- rep("7", nrow(cohort7_weight))


```
Weight data from cohort 7 and 8 were read in from `r cohort7_file` and `r cohort8_file`.

```{r process_cohort8, message=FALSE, echo=FALSE, warning=FALSE}

cohort8_file <- "../data/raw/Cohort_8_Raw_Data.csv"
cohort8.data <- read.csv(cohort8_file, row.names="X")

starting_date <- '2014-11-26'
date_before_fasting <- '2014-12-01'
date_after_fasting <- '2014-12-02'
date_12weeks_MRI <- '2015-02-24'

date_of_interest <- c('2014-12-01', '2014-12-02', '2015-02-24')

sub_cohort8.data <- cohort8.data[cohort8.data$experiment.date %in% date_of_interest,]
#Rename the fasting stage based on the date
sub_cohort8.data$Fasting_stat <- ifelse(sub_cohort8.data$experiment.date=="2014-12-01", "Pre_fasted",ifelse(sub_cohort8.data$experiment.date=="2014-12-02","Fasted", "After_HFD"))
sub_cohort8.data$animal.MouseID <- as.factor(sub_cohort8.data$animal.MouseID)

#import body composition data
cohort8_weight <-dcast(sub_cohort8.data, animal.MouseID+Fasting_stat~assay.assay, value.var='values')
cohort8_weight <- cohort8_weight[ , -which(colnames(cohort8_weight)=="Grip Strength (4 Paw)")]

cohort8_weight$Lean <- as.numeric(cohort8_weight$'Lean Mass')/1000
cohort8_weight$Fat <- as.numeric(cohort8_weight$'Total Fat Mass')/1000
cohort8_weight$Weight <- as.numeric(cohort8_weight$'Body Weight')/1000
#calculate percent body fat
cohort8_weight$Percent.Fat <- cohort8_weight$Fat/cohort8_weight$Weight*100
#make animal id a factor
cohort8_weight$Subject <- as.factor(cohort8_weight$animal.MouseID)
cohort8_weight$Fasting_stat <- as.factor(cohort8_weight$Fasting_stat)
#set Pre_fasting as the reference level
cohort8_weight$Fasting_stat <- relevel(as.factor(cohort8_weight$Fasting_stat), ref='Pre_fasted')
cohort8_weight$Cohort <- rep("8", nrow(cohort8_weight))

##### combine cohort 7 and 8 data #####
cohort7_8 <- rbind(cohort7_weight, cohort8_weight)
cohort7.8 <- cohort7_8[,c(1:2, 6:11)]

#### combine cohort 5,6,8, and 9####
combined_weight <- rbind(cohort5.6, cohort7.8)
combined_weight$Fasting_stat <- factor(combined_weight$Fasting_stat, levels=c("Pre_fasted", "Fasted", "After_HFD"))
combined_weight$Cohort <- factor(combined_weight$Cohort)

se <- function (x) sd(x, na.rm=T)/sqrt(length(x))
weight.stat <- ddply(combined_weight, .(Cohort, Fasting_stat), summarise, mean = mean(Weight), se=se(Weight))

```

```{r bar_graphs,  echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(weight.stat, aes(x=Fasting_stat, y=mean, fill=Cohort)) +
    geom_bar(position=position_dodge(), stat="identity") +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se),width=.2,                    # Width of the error bars
  position=position_dodge(.9))+
    ylab("Weight (g)")
ggsave("figure/Body_weight_Barplot.pdf")

#Boxplot
ggplot(combined_weight, aes(x=Fasting_stat, y=Weight, color=Cohort, fill=Cohort)) +
  geom_point(position=position_jitterdodge(dodge.width=0.9)) +
      geom_boxplot( outlier.colour = NA, 
                        position = position_dodge(width=0.9)) +
      scale_fill_brewer(palette="Pastel2")
ggsave("figure/Body_weight_Boxplot.pdf")

```

#Tukey post-hoc analysis of body weight
```{r Analysis_of_Variance, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
library(xtable)
anova.init.weight <- aov(Weight~Cohort ,data=combined_weight[combined_weight$Fasting_stat=="Pre_fasted",])
layout(matrix(c(1,2,3,4),2,2)) # optional layout 
plot(anova.init.weight) # diagnostic plots
#summary(anova.init.weight)
pairwise.init.weight <- TukeyHSD(anova.init.weight)

print("Tukey's post-hoc pairwise comparisons of Pre_fasted weight", comment=FALSE)
print(xtable(pairwise.init.weight$Cohort), comment=FALSE)

anova.fasted.weight <- aov(Weight~Cohort ,data=combined_weight[combined_weight$Fasting_stat=="Fasted",])
layout(matrix(c(1,2,3,4),2,2)) # optional layout 
plot(anova.fasted.weight) # diagnostic plots
#summary(anova.fasted.weight)
pairwise.fasted.weight <- TukeyHSD(anova.fasted.weight)

print("Tukey's post-hoc pairwise comparisons of Fasted weight", comment=FALSE)
print(xtable(pairwise.fasted.weight$Cohort), comment=FALSE)

anova.hfd.weight <- aov(Weight~Cohort ,data=combined_weight[combined_weight$Fasting_stat=="After_HFD",])
layout(matrix(c(1,2,3,4),2,2)) # optional layout 
plot(anova.hfd.weight) # diagnostic plots
#summary(anova.hfd.weight)
pairwise.hfd.weight <- TukeyHSD(anova.hfd.weight)

print("Tukey's post-hoc pairwise comparisons of Fasted weight", comment=FALSE)
print(xtable(pairwise.hfd.weight$Cohort), comment=FALSE)

```

```{r histograms,  echo=FALSE, message=FALSE, warning=FALSE}
#histogram
library(sm)
par(mfrow=c(2,2))
sm.density.compare(combined_weight$Weight[combined_weight$Fasting_stat=="Pre_fasted"], combined_weight$Cohort[combined_weight$Fasting_stat=="Pre_fasted"], xlab="Pre-fasted weight")
legend("topright", levels(combined_weight$Cohort), fill=2+(0:nlevels(combined_weight$Cohort)))
ggsave("figure/Density_Pre_fasted.pdf")

sm.density.compare(combined_weight$Weight[combined_weight$Fasting_stat=="Fasted"], combined_weight$Cohort[combined_weight$Fasting_stat=="Fasted"], xlab="Fasted Body weight")
legend("topright", levels(combined_weight$Cohort), fill=2+(0:nlevels(combined_weight$Cohort)))
ggsave("figure/Density_Fasted.pdf")

sm.density.compare(combined_weight$Weight[combined_weight$Fasting_stat=="After_HFD"], combined_weight$Cohort[combined_weight$Fasting_stat=="After_HFD"], xlab="Body weight after HFD")
legend("topright", levels(combined_weight$Cohort), fill=2+(0:nlevels(combined_weight$Cohort)))
ggsave("figure/Density_After_HFD.pdf")


#ggplot(combined_weight, aes(x=Weight,  fill=Cohort)) + 
 # geom_histogram(binwidth = 0.05) + theme(legend.position="top")+
 # facet_grid(Cohort~Fasting_stat) + 
 # ggtitle("Density of Body weight")

```
 
  Overall, body weight appeared to be normally distributed across different feeding stages, except for cohort 5. This may be due to one or two mice with outlier weights. 

#Checking the normality of weight loss and weight gain
```{r WL_vs_WG, message=FALSE, echo=FALSE, warning=FALSE}
weight.data <- combined_weight[, c(1:2, 5, 8)]
weight.wide <- reshape(combined_weight, 
  timevar = "Fasting_stat",
  idvar = c( "animal.MouseID", "Cohort", "Subject"),
  direction = "wide")

weight.wide$WL <- weight.wide$Weight.Pre_fasted - weight.wide$Weight.Fasted
weight.wide$WG <- weight.wide$Weight.After_HFD - weight.wide$Weight.Fasted

weight.wide$WL.pct <- (weight.wide$Weight.Pre_fasted - weight.wide$Weight.Fasted)*100/weight.wide$Weight.Pre_fasted
weight.wide$WG.pct <- (weight.wide$Weight.After_HFD - weight.wide$Weight.Fasted)*100/weight.wide$Weight.Fasted

#only select those with WL.pct >=0
weight.wide <- weight.wide[weight.wide$WL.pct >=0,]

par(mfrow=c(2,2))
qqnorm(weight.wide$WL, main="Absolute WL") 
qqline(weight.wide$WL)
qqnorm(weight.wide$WG, main="Absolute WG")
qqline(weight.wide$WG)
qqnorm(weight.wide$WL.pct, main="Percent WL")
qqline(weight.wide$WL.pct)
qqnorm(weight.wide$WG.pct, main="Percent WG")
qqline(weight.wide$WG.pct)

hist(weight.wide$WL, breaks = 15, freq = F, xlab = 'Weight Loss(g)',  ylab = 'Probability', main = 'Histogram of WL with Density Plot')
lines(density(weight.wide$WL, na.rm = T, from = min(weight.wide$WL), to = max(weight.wide$WL)))
hist(weight.wide$WG, breaks = 15, freq = F, xlab = 'Weight Loss(g)',  ylab = 'Probability', main = 'Histogram of WG with Density Plot')
lines(density(weight.wide$WG, na.rm = T, from = min(weight.wide$WG), to = max(weight.wide$WG)))
hist(weight.wide$WL.pct, breaks = 15, freq = F, xlab = 'Weight Loss(%)',  ylab = 'Probability', main = 'Histogram of %WL with Density Plot')
lines(density(weight.wide$WL.pct, na.rm = T, from = min(weight.wide$WL.pct), to = max(weight.wide$WL.pct)))
hist(weight.wide$WG.pct, breaks = 15, freq = F, xlab = 'Weight Gain(%)',  ylab = 'Probability', main = 'Histogram of %WG with Density Plot')
lines(density(weight.wide$WG.pct, na.rm = T, from = min(weight.wide$WG.pct), to = max(weight.wide$WG.pct)))

p.WL <- shapiro.test(weight.wide$WL)
p.WG <- shapiro.test(weight.wide$WG)
```
  
  WL and WG are not normally distributed as shown in the QQ plots, density plots and histograms. Shapiro p-values for WL and WG are `r p.WL$p.value` and `r p.WG$p.value`, respectively. Therefore, Spearman's rho or Kendall's tau correlation coefficient will be computed.


```{r create_plot_function, message=FALSE, echo=FALSE, warning=FALSE}
### create a function that draws a scatterplot between the 'xaxis' and 'yaxis' variables and show 'corr.type' correlation and the p-values
## data= data set
## xaxis= variable to be in the axis
## yaxis = variable to be yaxis
## col = variable used to color the dots or the lines
## title = title of the graph
## font.x = x coordinator for the correlation output
## font.y = y coordinator for the correlation output
## font.size = size of the font for correlation result output
## corr.type = type of correlation to be used, e.g. Kendall, Pearson, Spearman
####
corrplot <- function (data, xaxis, yaxis, col, title, font.x, font.y, font.size, corr.type){
  data$xaxis <- data[,xaxis]
  data$yaxis <- data[,yaxis]
  data$col <- data[, col]
  corSet <- round(cor(data$xaxis, data$yaxis, method=corr.type),3)
  p.cor <- signif(cor.test(data$xaxis, data$yaxis, method=corr.type, exact=FALSE)$p.value,digits=3)
  data$corSet <- corSet
  data$p.cor <- p.cor
  #print(p.cor)
  p <- ggplot(data, aes(x=xaxis, y=yaxis))
  p + geom_point() + theme(legend.position="top")+
  geom_smooth(method=lm, se=T) + geom_text(aes(label=paste("r=", corSet, sep="")), x=font.x, y=font.y, size=font.size) +
  geom_text(aes(label=paste("p-value=", p.cor, sep="")), x=font.x, y=font.y-3, size=font.size) +
  xlab(xaxis) + ylab(yaxis) +
  ggtitle(title)
  #ggsave(filename)
}

### create a function that draws multiple scatterplots between the 'xaxis' and 'yaxis' variables and show Kendall's correlation and the p-values ####
mult.corrplot <- function (data, xaxis, yaxis, col, title, font.x, font.y, font.size, corr.type){
  data$xaxis <- data[,xaxis]
  data$yaxis <- data[,yaxis]
  data$col <- data[, col]
  #data$corr.type <- corr.type
  #print(data$corr.type)
  #print(dum.corr.type)
  corSet <- ddply(data, .(col), summarise, cor=round(cor(xaxis, yaxis, method="kendall"),3))
  #print(corSet)
  p.cor <- ddply(data, .(col), summarise, p=signif(cor.test(xaxis, yaxis, method="kendall", exact=FALSE)$p.value,digits=3))
  #print(p.cor)
  p <- ggplot(data, aes(x=xaxis, y=yaxis, fill=col, color=col))
  p + geom_point() + facet_grid(col~.)+
      theme(legend.position="top")+
      geom_smooth(method=lm, se=T) + geom_text(data=corSet, aes(label=paste("r=", cor, sep="")), x=font.x, y=font.y, size=font.size) +
      geom_text(data=p.cor, aes(label=paste("p-value=", p, sep="")), x=font.x, y=font.y-10, size=font.size) +
      xlab(xaxis) + ylab(yaxis) +
      ggtitle(title)
}

```

Relationship between Weight Loss after fasting and Weight Gain after HFD
============
```{r WLfasted_vs_WG, message=FALSE, echo=FALSE, warning=FALSE}
par(mfrow=c(1,1))
##### Absolute weight ##########
ggplot(weight.wide, aes(x=WL, y=WG, color=Cohort, fill=Cohort)) +
  geom_point() + theme(legend.position="top")+
  geom_smooth(method=lm, se=T)
ggsave("figure/Correlation_WL_WG.pdf")


mult.corrplot(weight.wide, "WL", "WG", "Cohort", "Correlation between WL and WG in each cohort",3, 25, 4, "kendall")
ggsave("figure/Correlation_WL_WG_each_cohort.pdf")

### all corhorts ###
corrplot(weight.wide, "WL", "WG", "Cohort", "All cohort combined (absolute values)", 3, 25, 5, "kendall")
ggsave("figure/Correlation_WL_WG_all_cohort.pdf")

########Percentage weight #########
corspct <- ddply(weight.wide, .(Cohort), summarise, cor=round(cor(WL.pct, WG.pct, method="kendall"),2))
pcorrPct <- ddply(weight.wide, .(Cohort), summarise, p=signif(cor.test(WL.pct, WG.pct, method="kendall")$p.value,digits=3))

mult.corrplot(weight.wide, "WL.pct", "WG.pct", "Cohort", "Correlation between %WL and %WG in each cohort", 2.5, 100, 4, "kendall")
ggsave("figure/Correlation_percentWL_percentWG_each_cohort.pdf")

### all corhorts ###
corpct <- round(cor(weight.wide$WL.pct, weight.wide$WG.pct, method="kendall"), 3)
pcorPct <- signif(cor.test(weight.wide$WL.pct, weight.wide$WG.pct, method="kendall")$p.value,digits=3)
corrplot(weight.wide, "WL.pct", "WG.pct", "Cohort", "All cohort combined (percentage)", 2.5, 120, 5, "kendall")
ggsave("figure/Correlation_percentWL_percentWG_all_cohort.pdf")

require(NSM3)
ktau.ci <- kendall.ci(weight.wide$WL, weight.wide$WG, B = 1000)

```
  
  There is a negative correlation between weight loss (or % weight loss) and weight gain (or %weight gain) after HFD in cohort 5 and 6 while the direction of the relationship reverse in cohort 7 and 8. Only the relationship in cohort 8 is significant after Benjamini Hochberg adjustment with p-adjust=0.01536. When combining all 4 cohorts together, there is a `r corpct` correlation between WL and WG and this correlation is statistically significant with p-value=`r pcorPct`. This is likely due to the increase in sample size from about 20 mice in each cohort to about 100 mice in all. A more reliable information to look at is the 95% CI for the Kendall Tau correlation coefficient. The `r print(kendall.ci(weight.wide$WL, weight.wide$WG, B = 1000))`. The range of this CI is reasonable, indicating that the effect may be legitimate.  

Relationship between Fat mass and HFD Weight
===============
```{r fatMass_vs_WG, message=FALSE, warning=FALSE, echo=FALSE}

#### % Fat mass at starting ######
sub.data <- weight.wide[weight.wide$Cohort%in%c('7', '8'),]
corsFatInit <- ddply(sub.data, .(Cohort), summarise, cor=round(cor(Percent.Fat.Pre_fasted, WG, method="spearman"),2))
pcorrFatInit <- ddply(sub.data, .(Cohort), summarise, p=signif(cor.test(Percent.Fat.Pre_fasted, WG, method="spearman", exact=FALSE)$p.value,digits=3))

mult.corrplot(sub.data, "Percent.Fat.Pre_fasted", "WG.pct", "Cohort", "Correlation between %Fat Mass initial and % WG in each cohort", 10, 100, 4, "kendall")
ggsave("figure/Correlation_PercentFatMassInit_PercentWG_each_cohort.pdf")

### Percent Fat Mass after fasting######

mult.corrplot(sub.data, "Percent.Fat.Fasted", "WG.pct", "Cohort", "Correlation between %Fat Mass fasted and WG in each cohort", 9, 100, 4, "kendall")
ggsave("figure/Correlation_PercentFatMassFasted_PercentWG_each_cohort.pdf")

####3 Combined cohort 7 and 8 ######
corr <- round(cor(sub.data$Percent.Fat.Fasted, sub.data$WG.pct, method="spearman"),3)
pcorr <- signif(cor.test(sub.data$Percent.Fat.Fasted, sub.data$WG.pct, method="spearman", exact=FALSE)$p.value,digits=3)
corrplot(sub.data, "Percent.Fat.Fasted", "WG.pct", "Cohort", "% Fat Mass Fasted vs Weight Gain", 9, 100, 5, "spearman")
ggsave("figure/Correlation_PercentFatFasted_PercentWG_cohort7_8.pdf")

```

  There is no correlation between percent fat mass after fasting and the absolute weight gain, r=`r corr` and p-value=`r pcorr`.
  
Session Information
---------------------
```{r sessionInfo, echo=F}
sessionInfo()
```