---
title: "Compare_cohorts"
author: "Quynh Tran"
date: "November 4, 2015"
output: html_document
---

This script was most recently run on `r date()`. This script is to compare the mice body weight across different cohorts. The script also explores the relationship between initial weight, weight loss and the weights after HFD.

```{r process_cohort5_6_data, message=FALSE, echo=FALSE, warning=FALSE}
previous_cohort_file <- "../data/processed/fasting_weights_file.csv"
prev_cohort_data <- read.csv(previous_cohort_file)

#only get HFD mice from previous cohorts
prev.data <- prev_cohort_data[prev_cohort_data$Treatment=="High Fat Diet",]
prev.data <- prev.data[!is.na(prev.data$X),]
prev.data$Subject <- substring(prev.data$Animal, 18,20)

prev.long <- reshape(prev.data, 
  varying = c("Early.Fed", "Early.Fasted", "Late.Fed", "Late.Fasted"), 
  v.names = "Weight",
  timevar = "Fasting_stat", 
  times = c("Early.Fed", "Early.Fasted", "Late.Fed", "Late.Fasted"), 
  new.row.names = 1:1000,
  direction = "long")
prev.long$Fasting_stat <- ifelse(prev.long$Fasting_stat=="Early.Fed", "Initial_weight", ifelse(prev.long$Fasting_stat=="Early.Fasted", "Fasted", ifelse(prev.long$Fasting_stat=="Late.Fed", "After_HFD", prev.long$Fasting_stat)))


sub.prev.long <- prev.long[prev.long$Fasting_stat%in% c("Initial_weight", "Fasted", "After_HFD"),]
sub.prev.long <- sub.prev.long[!is.na(sub.prev.long$X),]

#create a data set for cohort 5 and 6 that has the same structure as cohort 8 and 9
cohort5.6 <- data.frame(animal.MouseID=sub.prev.long$Subject, Fasting_stat=sub.prev.long$Fasting_stat, Lean = numeric(nrow(sub.prev.long)), Fat=numeric(nrow(sub.prev.long)), Weight=sub.prev.long$Weight, Percent.Fat=numeric(nrow(sub.prev.long)), Subject=sub.prev.long$Subject, Cohort=substring(sub.prev.long$Cohort, 8,9))


```
Weight data from cohort 5 and 6 were read in from `r previous_cohort_file`.

```{r process_cohort7, message=FALSE, echo=FALSE, warning=FALSE}
cohort8_file <- "../data/raw/Cohort_8_Raw_Data.csv"
cohort9_file <- "../data/raw/Cohort_9_Raw_Data.csv"

cohort8.data <- read.csv(cohort8_file, row.names="X")
cohort9.data <- read.csv(cohort9_file, row.names="X")

starting_date <- '2014-10-17'
date_after_fasting <- '2014-10-21'
date_before_fasting <- '2014-10-20'
date_12weeks_MRI <- '2015-01-16'


date_of_interest <- c('2014-10-17', '2014-10-21', '2014-10-20', '2015-01-16')

sub_cohort8.data <- cohort8.data[cohort8.data$experiment.date %in% date_of_interest,]
#Rename the fasting stage based on the date
sub_cohort8.data$Fasting_stat <- ifelse(sub_cohort8.data$experiment.date=="2014-10-20", "Pre_fasted",ifelse(sub_cohort8.data$experiment.date=="2014-10-21","Fasted",ifelse(sub_cohort8.data$experiment.date=="2014-10-17", "Initial_weight", "After_HFD")))
sub_cohort8.data$animal.MouseID <- as.factor(sub_cohort8.data$animal.MouseID)

#import body composition data
library(reshape2)
library(plyr)
#there should be Lean Mass for this record, but some how it was labeled "Total Fat Mass"
sub_cohort8.data$assay.assay[which(sub_cohort8.data$assay.assay=="Total Fat Mass" & sub_cohort8.data$values=="22200")] <- "Lean Mass"

cohort8_weight <-dcast(sub_cohort8.data, animal.MouseID+Fasting_stat~assay.assay, value.var='values')

cohort8_weight$Lean <- as.numeric(cohort8_weight$'Lean Mass')/1000
cohort8_weight$Fat <- as.numeric(cohort8_weight$'Total Fat Mass')/1000
cohort8_weight$Weight <- as.numeric(cohort8_weight$'Body Weight')/1000
#calculate percent body fat
cohort8_weight$Percent.Fat <- cohort8_weight$Fat/cohort8_weight$Weight*100
#make animal id a factor
cohort8_weight$Subject <- as.factor(cohort8_weight$animal.MouseID)
cohort8_weight$Fasting_stat <- as.factor(cohort8_weight$Fasting_stat)
#set Pre_fasting as the reference level
cohort8_weight$Fasting_stat <- relevel(as.factor(cohort8_weight$Fasting_stat), ref='Pre_fasted')
cohort8_weight$Cohort <- rep("7", nrow(cohort8_weight))


```
Weight data from cohort 8 and 9 were read in from `r cohort8_file` and `r cohort9_file`.

```{r process_cohort8, message=FALSE, echo=FALSE, warning=FALSE}

cohort9_file <- "../data/raw/Cohort_9_Raw_Data.csv"
cohort9.data <- read.csv(cohort9_file, row.names="X")

starting_date <- '2014-11-26'
date_before_fasting <- '2014-12-01'
date_after_fasting <- '2014-12-02'
date_12weeks_MRI <- '2015-02-24'

date_of_interest <- c('2014-11-26', '2014-12-01', '2014-12-02', '2015-02-24')

sub_cohort9.data <- cohort9.data[cohort9.data$experiment.date %in% date_of_interest,]
#Rename the fasting stage based on the date
sub_cohort9.data$Fasting_stat <- ifelse(sub_cohort9.data$experiment.date=="2014-12-01", "Pre_fasted",ifelse(sub_cohort9.data$experiment.date=="2014-12-02","Fasted",ifelse(sub_cohort9.data$experiment.date=="2014-11-26", "Initial_weight", "After_HFD")))
sub_cohort9.data$animal.MouseID <- as.factor(sub_cohort9.data$animal.MouseID)

#import body composition data
cohort9_weight <-dcast(sub_cohort9.data, animal.MouseID+Fasting_stat~assay.assay, value.var='values')
cohort9_weight <- cohort9_weight[ , -which(colnames(cohort9_weight)=="Grip Strength (4 Paw)")]

cohort9_weight$Lean <- as.numeric(cohort9_weight$'Lean Mass')/1000
cohort9_weight$Fat <- as.numeric(cohort9_weight$'Total Fat Mass')/1000
cohort9_weight$Weight <- as.numeric(cohort9_weight$'Body Weight')/1000
#calculate percent body fat
cohort9_weight$Percent.Fat <- cohort9_weight$Fat/cohort9_weight$Weight*100
#make animal id a factor
cohort9_weight$Subject <- as.factor(cohort9_weight$animal.MouseID)
cohort9_weight$Fasting_stat <- as.factor(cohort9_weight$Fasting_stat)
#set Pre_fasting as the reference level
cohort9_weight$Fasting_stat <- relevel(as.factor(cohort9_weight$Fasting_stat), ref='Pre_fasted')
cohort9_weight$Cohort <- rep("8", nrow(cohort9_weight))

##### combine cohort 7 and 8 data #####
cohort8_9 <- rbind(cohort8_weight, cohort9_weight)
cohort8.9 <- cohort8_9[,c(1:2, 6:11)]
#### combine cohort 5,6,8, and 9####
combined_weight <- rbind(cohort5.6, cohort8.9)
combined_weight$Fasting_stat <- factor(combined_weight$Fasting_stat, levels=c("Initial_weight", "Pre_fasted", "Fasted", "After_HFD"))
combined_weight$Cohort <- factor(combined_weight$Cohort)

se <- function (x) sd(x, na.rm=T)/sqrt(length(x))
weight.stat <- ddply(combined_weight, .(Cohort, Fasting_stat), summarise, mean = mean(Weight), se=se(Weight))

```

```{r bar_graphs,  echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(weight.stat, aes(x=Fasting_stat, y=mean, fill=Cohort)) +
    geom_bar(position=position_dodge(), stat="identity") +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se),width=.2,                    # Width of the error bars
  position=position_dodge(.9))+
    ylab("Weight (g)")
ggsave("figure/Body_weight_Barplot.pdf")

#Boxplot
ggplot(combined_weight, aes(x=Fasting_stat, y=Weight, color=Cohort, fill=Cohort)) +
  geom_point(position=position_jitterdodge(dodge.width=0.9)) +
      geom_boxplot( outlier.colour = NA, 
                        position = position_dodge(width=0.9)) +
      scale_fill_brewer(palette="Pastel2")
ggsave("figure/Body_weight_Boxplot.pdf")

```

#Tukey post-hoc analysis of body weight
```{r Analysis_of_Variance, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
library(xtable)
anova.init.weight <- aov(Weight~Cohort ,data=combined_weight[combined_weight$Fasting_stat=="Initial_weight",])
layout(matrix(c(1,2,3,4),2,2)) # optional layout 
plot(anova.init.weight) # diagnostic plots
#summary(anova.init.weight)
pairwise.init.weight <- TukeyHSD(anova.init.weight)

print("Tukey's post-hoc pairwise comparisons of Ininital weight", comment=FALSE)
print(xtable(pairwise.init.weight$Cohort), comment=FALSE)

anova.fasted.weight <- aov(Weight~Cohort ,data=combined_weight[combined_weight$Fasting_stat=="Fasted",])
layout(matrix(c(1,2,3,4),2,2)) # optional layout 
plot(anova.fasted.weight) # diagnostic plots
#summary(anova.fasted.weight)
pairwise.fasted.weight <- TukeyHSD(anova.fasted.weight)

print("Tukey's post-hoc pairwise comparisons of Fasted weight", comment=FALSE)
print(xtable(pairwise.fasted.weight$Cohort), comment=FALSE)

anova.hfd.weight <- aov(Weight~Cohort ,data=combined_weight[combined_weight$Fasting_stat=="After_HFD",])
layout(matrix(c(1,2,3,4),2,2)) # optional layout 
plot(anova.hfd.weight) # diagnostic plots
#summary(anova.hfd.weight)
pairwise.hfd.weight <- TukeyHSD(anova.hfd.weight)

print("Tukey's post-hoc pairwise comparisons of Fasted weight", comment=FALSE)
print(xtable(pairwise.hfd.weight$Cohort), comment=FALSE)

```

```{r histograms,  echo=FALSE, message=FALSE, warning=FALSE}
#histogram
library(sm)
par(mfrow=c(2,2))
sm.density.compare(combined_weight$Weight[combined_weight$Fasting_stat=="Initial_weight"], combined_weight$Cohort[combined_weight$Fasting_stat=="Initial_weight"], xlab="Initial Body weight")
legend("topright", levels(combined_weight$Cohort), fill=2+(0:nlevels(combined_weight$Cohort)))
ggsave("figure/Density_Initial_Weight.pdf")

sm.density.compare(combined_weight$Weight[combined_weight$Fasting_stat=="Fasted"], combined_weight$Cohort[combined_weight$Fasting_stat=="Fasted"], xlab="Fasted Body weight")
legend("topright", levels(combined_weight$Cohort), fill=2+(0:nlevels(combined_weight$Cohort)))
ggsave("figure/Density_Fasted.pdf")

sm.density.compare(combined_weight$Weight[combined_weight$Fasting_stat=="After_HFD"], combined_weight$Cohort[combined_weight$Fasting_stat=="After_HFD"], xlab="Body weight after HFD")
legend("topright", levels(combined_weight$Cohort), fill=2+(0:nlevels(combined_weight$Cohort)))
ggsave("figure/Density_After_HFD.pdf")


#ggplot(combined_weight, aes(x=Weight,  fill=Cohort)) + 
 # geom_histogram(binwidth = 0.05) + theme(legend.position="top")+
 # facet_grid(Cohort~Fasting_stat) + 
 # ggtitle("Density of Body weight")

```
 
  Overall, body weight appeared to be normally distributed across different feeding stages, except for cohort 5. This may be due to one or two mice with outlier weights. 

#Checking the normality of weight loss and weight gain
```{r WL_vs_WG, message=FALSE, echo=FALSE, warning=FALSE}
weight.data <- combined_weight[, c(1:2, 5, 8)]
weight.wide <- reshape(combined_weight, 
  timevar = "Fasting_stat",
  idvar = c( "animal.MouseID", "Cohort", "Subject"),
  direction = "wide")

weight.wide$WL <- weight.wide$Weight.Initial_weight - weight.wide$Weight.Fasted
weight.wide$WG <- weight.wide$Weight.After_HFD - weight.wide$Weight.Fasted

weight.wide$WL.pct <- (weight.wide$Weight.Initial_weight - weight.wide$Weight.Fasted)*100/weight.wide$Weight.Initial_weight
weight.wide$WG.pct <- (weight.wide$Weight.After_HFD - weight.wide$Weight.Fasted)*100/weight.wide$Weight.Fasted

par(mfrow=c(2,2))
qqnorm(weight.wide$WL, main="Absolute WL") 
qqline(weight.wide$WL)
qqnorm(weight.wide$WG, main="Absolute WG")
qqline(weight.wide$WG)
qqnorm(weight.wide$WL.pct, main="Percent WL")
qqline(weight.wide$WL.pct)
qqnorm(weight.wide$WG.pct, main="Percent WG")
qqline(weight.wide$WG.pct)

hist(weight.wide$WL, breaks = 15, freq = F, xlab = 'Weight Loss(g)',  ylab = 'Probability', main = 'Histogram of WL with Density Plot')
lines(density(weight.wide$WL, na.rm = T, from = min(weight.wide$WL), to = max(weight.wide$WL)))
hist(weight.wide$WG, breaks = 15, freq = F, xlab = 'Weight Loss(g)',  ylab = 'Probability', main = 'Histogram of WG with Density Plot')
lines(density(weight.wide$WG, na.rm = T, from = min(weight.wide$WG), to = max(weight.wide$WG)))
hist(weight.wide$WL.pct, breaks = 15, freq = F, xlab = 'Weight Loss(%)',  ylab = 'Probability', main = 'Histogram of %WL with Density Plot')
lines(density(weight.wide$WL.pct, na.rm = T, from = min(weight.wide$WL.pct), to = max(weight.wide$WL.pct)))
hist(weight.wide$WG.pct, breaks = 15, freq = F, xlab = 'Weight Gain(%)',  ylab = 'Probability', main = 'Histogram of %WG with Density Plot')
lines(density(weight.wide$WG.pct, na.rm = T, from = min(weight.wide$WG.pct), to = max(weight.wide$WG.pct)))

p.WL <- shapiro.test(weight.wide$WL)
p.WG <- shapiro.test(weight.wide$WG)
```
  
  WL and WG are not normally distributed as shown in the QQ plots, density plots and histograms. Shapiro p-values for WL and WG are `r p.WL$p.value` and `r p.WG$p.value`, respectively. Therefore, Spearman's rho or Kendall's tau correlation coefficient will be computed.


```{r create_plot_function, message=FALSE, echo=FALSE, warning=FALSE}
### create a function that draws a scatterplot between the 'xaxis' and 'yaxis' variables and show 'corr.type' correlation and the p-values ####
corrplot <- function (data, xaxis, yaxis, col, title, font.x, font.y, font.size, corr.type){
  data$xaxis <- data[,xaxis]
  data$yaxis <- data[,yaxis]
  data$col <- data[, col]
  corSet <- round(cor(data$xaxis, data$yaxis, method=corr.type),3)
  p.cor <- signif(cor.test(data$xaxis, data$yaxis, method=corr.type, exact=FALSE)$p.value,digits=3)
  data$corSet <- corSet
  data$p.cor <- p.cor
  #print(p.cor)
  p <- ggplot(data, aes(x=xaxis, y=yaxis))
  p + geom_point() + theme(legend.position="top")+
  geom_smooth(method=lm, se=T) + geom_text(aes(label=paste("r=", corSet, sep="")), x=font.x, y=font.y, size=font.size) +
  geom_text(aes(label=paste("p-value=", p.cor, sep="")), x=font.x, y=font.y-3, size=font.size) +
  xlab(xaxis) + ylab(yaxis) +
  ggtitle(title)
  #ggsave(filename)
}

### create a function that draws multiple scatterplots between the 'xaxis' and 'yaxis' variables and show Kendall's correlation and the p-values ####
mult.corrplot <- function (data, xaxis, yaxis, col, title, font.x, font.y, font.size, corr.type){
  data$xaxis <- data[,xaxis]
  data$yaxis <- data[,yaxis]
  data$col <- data[, col]
  #data$corr.type <- corr.type
  #print(data$corr.type)
  #print(dum.corr.type)
  corSet <- ddply(data, .(col), summarise, cor=round(cor(xaxis, yaxis, method="kendall"),3))
  #print(corSet)
  p.cor <- ddply(data, .(col), summarise, p=signif(cor.test(xaxis, yaxis, method="kendall", exact=FALSE)$p.value,digits=3))
  #print(p.cor)
  p <- ggplot(data, aes(x=xaxis, y=yaxis, fill=col, color=col))
  p + geom_point() + facet_grid(col~.)+
      theme(legend.position="top")+
      geom_smooth(method=lm, se=T) + geom_text(data=corSet, aes(label=paste("r=", cor, sep="")), x=font.x, y=font.y, size=font.size) +
      geom_text(data=p.cor, aes(label=paste("p-value=", p, sep="")), x=font.x, y=font.y-10, size=font.size) +
      xlab(xaxis) + ylab(yaxis) +
      ggtitle(title)
}

```

Relationship between Weight Loss after fasting and Weight Gain after HFD
============
```{r WLfasted_vs_WG, message=FALSE, echo=FALSE, warning=FALSE}
par(mfrow=c(1,1))
##### Absolute weight ##########
ggplot(weight.wide, aes(x=WL, y=WG, color=Cohort, fill=Cohort)) +
  geom_point() + theme(legend.position="top")+
  geom_smooth(method=lm, se=T)
ggsave("figure/Correlation_WL_WG.pdf")


mult.corrplot(weight.wide, "WL", "WG", "Cohort", "Correlation between WL and WG in each cohort", 5, 25, 4, "kendall")
ggsave("figure/Correlation_WL_WG_each_cohort.pdf")

### all corhorts ###
corrplot(weight.wide, "WL", "WG", "Cohort", "All cohort combined (absolute values)", 5, 25, 5, "kendall")
ggsave("figure/Correlation_WL_WG_all_cohort.pdf")

########Percentage weight #########
corspct <- ddply(weight.wide, .(Cohort), summarise, cor=round(cor(WL.pct, WG.pct, method="kendall"),2))
pcorrPct <- ddply(weight.wide, .(Cohort), summarise, p=signif(cor.test(WL.pct, WG.pct, method="kendall")$p.value,digits=3))

mult.corrplot(weight.wide, "WL.pct", "WG.pct", "Cohort", "Correlation between %WL and %WG in each cohort", 20, 120, 4, "kendall")
ggsave("figure/Correlation_percentWL_percentWG_each_cohort.pdf")

### all corhorts ###
corpct <- round(cor(weight.wide$WL.pct, weight.wide$WG.pct, method="kendall"), 3)
pcorPct <- signif(cor.test(weight.wide$WL.pct, weight.wide$WG.pct, method="kendall")$p.value,digits=3)
corrplot(weight.wide, "WL.pct", "WG.pct", "Cohort", "All cohort combined (percentage)", 20, 120, 5, "kendall")
ggsave("figure/Correlation_percentWL_percentWG_all_cohort.pdf")

require(NSM3)
ktau.ci <- kendall.ci(weight.wide$WL, weight.wide$WG, B = 1000)

```
  
  There is a negative correlation between weight loss (or % weight loss) and weight gain (or %weight gain) after HFD in cohort 5 and 6 while the direction of the relationship reverse in cohort 8 and 9. Only the relationship in cohort 9 is significant after Benjamini Hochberg adjustment with p-adjust=0.01536. When combining all 4 cohorts together, there is a `r corpct` correlation between WL and WG and this correlation is statistically significant with p-value=`r pcorPct`. This is likely due to the increase in sample size from about 20 mice in each cohort to about 100 mice in all. A more reliable information to look at is the 95% CI for the spearman correlation coefficient. The `r print(kendall.ci(weight.wide$WL, weight.wide$WG, B = 1000))`. The range of this CI is wide, indicating that we don't have much knowledge about the effect and more information or more murine experiments are needed.  


Relationship between Initial Weight and HFD Weight
===============
```{r initWeight_vs_WG, message=FALSE, echo=FALSE, warning=FALSE}
mult.corrplot(weight.wide, "Weight.Initial_weight", "WG", "Cohort", "Correlation between Initial weight and WG in each cohort", 29, 15, 4, "kendall")
ggsave("figure/Correlation_InitWeight_WG_each_cohort.pdf")

## All cohort combined ####
corrplot(weight.wide, "Weight.Initial_weight", "WG", "Cohort", "Initial Weight vs Weight Gain", 29, 15, 5, "spearman")
ggsave("figure/Correlation_InitWeight_WG_all_cohort.pdf")

require(RVAideMemoire)
spm.ci <- spearman.ci(weight.wide$Weight.Initial_weight, weight.wide$WG, nrep = 1000, conf.level = 0.95)
kendall.ci(weight.wide$Weight.Initial_weight, weight.wide$WG, B = 1000)
```
  
  There is no correlation between initial weight and weight gain after HFD in any individual cohort. All p-values are greater than 0.05, suggesting that we cannot reject the null hypothesis that there is no correlation between init weight and weight gain. When combining all 4 cohorts together, there is a significant correlation between initial weight and weight gain after HFD. This is likely due to the increase in sample size from about 20 mice in each cohort to about 100 mice in all. A more reliable information to look at is the 95% CI for the Spearman correlation coefficient. The 95% CI is (`r spm.ci$conf.int[[1]]`, `r spm.ci$conf.int[[2]]`). The range of this CI is wide, indicating that we don't have much knowledge about the effect and more information or more murine experiment is needed.

Relationship between Fat mass and HFD Weight
===============
```{r fatMass_vs_WG, message=FALSE, warning=FALSE, echo=FALSE}

#### % Fat mass at starting ######
sub.data <- weight.wide[weight.wide$Cohort%in%c('7', '8'),]
corsFatInit <- ddply(sub.data, .(Cohort), summarise, cor=round(cor(Percent.Fat.Initial_weight, WG, method="spearman"),2))
pcorrFatInit <- ddply(sub.data, .(Cohort), summarise, p=signif(cor.test(Percent.Fat.Initial_weight, WG, method="spearman", exact=FALSE)$p.value,digits=3))

mult.corrplot(sub.data, "Percent.Fat.Initial_weight", "WG.pct", "Cohort", "Correlation between %Fat Mass initial and % WG in each cohort", 13, 100, 4, "kendall")
ggsave("figure/Correlation_PercentFatMassInit_PercentWG_each_cohort.pdf")

### Percent Fat Mass after fasting######

mult.corrplot(sub.data, "Percent.Fat.Fasted", "WG.pct", "Cohort", "Correlation between %Fat Mass fasted and WG in each cohort", 9, 100, 4, "kendall")
ggsave("figure/Correlation_PercentFatMassFasted_PercentWG_each_cohort.pdf")

####3 Combined cohort 7 and 8 ######
corr <- round(cor(sub.data$Percent.Fat.Fasted, sub.data$WG.pct, method="spearman"),3)
pcorr <- signif(cor.test(sub.data$Percent.Fat.Fasted, sub.data$WG.pct, method="spearman", exact=FALSE)$p.value,digits=3)
corrplot(sub.data, "Percent.Fat.Fasted", "WG.pct", "Cohort", "% Fat Mass Fasted vs Weight Gain", 9, 100, 5, "spearman")
ggsave("figure/Correlation_PercentFatFasted_PercentWG_cohort8_9.pdf")

```

  There is no correlation between percent fat mass after fasting and the absolute weight gain, r=`r corr` and p-value=`r pcorr`.
  
Session Information
---------------------
```{r sessionInfo, echo=F}
sessionInfo()
```